### 2017-03-26 slides/athenatalk.tex
[ slides ] .tex skeleton added.

\documentclass[hyperref={pdfpagelabels=false}]{beamer}
\beamertemplatenavigationsymbolsempty
\usepackage[utf8x]{inputenc}
\newcommand{\textdf}[1]{\textbf{\textsf{#1}}\index{#1}}
\newcommand{\agda}[0]{\textbf{\texttt{Agda}\,}}
\newcommand{\keyword}[1]{\textit{#1}}
\newcommand{\miss}[0]{{\colorbox{amber}{\textdf{?}}}}
\newcommand{\ttgray}[1]{\texttt{\color{gray} #1}}
\newcommand{\shell}[1]{\texttt{\color{gray} #1}}
\newcommand{\highlight}[1]{\texttt{\color{bondiblue} #1}}
\let\Tiny=\tiny
\usepackage{lmodern}
\usepackage{url}
\usepackage{bussproofs}
\definecolor{amethyst}{rgb}{0.6, 0.4, 0.8}
\definecolor{antiquefuchsia}{rgb}{0.57, 0.36, 0.51}
\definecolor{ballblue}{rgb}{0.13, 0.67, 0.8}
\definecolor{blue(munsell)}{rgb}{0.0, 0.5, 0.69}
\definecolor{blue(pigment)}{rgb}{0.2, 0.2, 0.6}
\definecolor{brightmaroon}{rgb}{0.76, 0.13, 0.28}
\definecolor{bondiblue}{rgb}{0.0, 0.58, 0.71}
\definecolor{cadet}{rgb}{0.33, 0.41, 0.47}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}

\hypersetup{linktocpage}
\hypersetup{
colorlinks,
citecolor=amethyst,
filecolor=black,
linkcolor=blue(munsell),
urlcolor=blue(munsell)
}

\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath, amsthm, amssymb}

\title{Proof Reconstruction with Athena\\(Work in Progress)}
% \date{\small{\jobname}}
\date{9th May}
\author{By Jonathan Prieto-Cubides}
\institute{
EAFIT University\\
Medell\'in, Colombia
}

\begin{document}

\setcounter{page}{1}
\maketitle

\begin{frame}{References}

\end{frame}

\end{document}