### 2017-04-24 slides/ideas.tex
[ slides ] diagrams added.

\begin{frame}{Keywords}

 \tikz[overlay,remember picture]
  \node[opacity=0.2]at (0.8\textwidth, 0.5cm)
    {\includegraphics[width=0.3\textwidth]{figures/metis}};

  \begin{block}{Proof Assistant}
    Is a software program that work together with a human in order to give formal proofs to problems in a wide range of topics.\\
    - Agda, Isabelle/HOL, HOL Light, Coq.
  \end{block}
  \begin{block}{\text{A}utomatic \textbf{T}heorem \textbf{P}rover}
    Is software program that tries to prove conjecture based on a list of hypothesis.\\
    - Metis, Z3, Vampire, EProver, iProver, SPSS.
  \end{block}
\end{frame}

\begin{frame}{Athena}{asdf}
  Is a Haskell program that translates proofs given by Metis ATP in TSTP format
  to Agda code.

  \begin{figure}
    \includegraphics[scale=0.85]{figures/diagram.pdf}
    \label{im:athena}
    \caption{Proof reconstruction work flow.}
  \end{figure}
\end{frame}

\begin{frame}{TPTP}
  Is a text format to encode problems in different theories. Its goal is provide
  a standard for all ATPs in the input file format. The version for derivations
  is named TSTP but this one is not a standard yet.\\[2mm]

  A TPTP file looks like:

  \problemtptp[basic-4.tptp]{../test/prop-pack/problems/basic/basic-4.tptp}{2}

\end{frame}

\begin{frame}{TSTP}
  A derivation of the previous problem output by Metis ATP.
{\footnotesize
  \problemtptp[basic-4.tstp]{../test/prop-pack/problems/basic/basic-4.tstp}{0}
}
\end{frame}

\begin{frame}{Agda}
   A verified proof of the previous derivation output by Athena.

  \begin{figure}
    \includegraphics[scale=0.35]{basic-4}
    \label{im:basic}
    \caption{Type-checked Proof.}
  \end{figure}

\end{frame}

\begin{frame}{Natural deduction tree}
  $\Gamma , p \vdash p$
\begin{proof}\hfill
\label{basic-der}
\begin{prooftree}
\AxiomC{$\Gamma \vdash a_1$}
\RightLabel{\scriptsize({\bfseries weaken $\neg\ \text{subgoal}_0$})}
\UnaryInfC{$\Gamma, \neg\ \text{subgoal}_0\,\vdash a_1$}
\RightLabel{\scriptsize({\bfseries canonicalize})}
\UnaryInfC{$\Gamma, \neg\ \text{subgoal}_0\,\vdash p$}
\AxiomC{$\Gamma \vdash p$}
\RightLabel{\scriptsize({\bfseries assume})}
\UnaryInfC{$\Gamma, \neg\ \text{subgoal}_0\,\vdash \neg p$}
\RightLabel{\scriptsize({\bfseries canonicalize})}
\UnaryInfC{$\Gamma, \neg\ \text{subgoal}_0\,\vdash \neg p$}
\RightLabel{\scriptsize({\bfseries simplify})}
\UnaryInfC{$\Gamma, \neg\ \text{subgoal}_0\,\vdash \neg p$}
\BinaryInfC{$\bot$}
\RightLabel{\scriptsize({\bfseries RAA})}
\UnaryInfC{$\Gamma \vdash p$}
\end{prooftree}
% A tree deduction of problem \ref{prob1}
\end{proof}
\end{frame}
% The tool : Athena.
% Its components:
% - agda-prop (logic framework)
% - agda-metis (ATP rules)
% - prop-pack (problems)

\begin{frame}{Future work}

  \begin{itemize}
  \item
    Extend the libraries Agda-Prop and Agda-Metis to handle predicated logic
  \item
    Reconstruct proofs of eprover in Agda-EProver
  \item
    Support more versions of GHC in Athena
  \end{itemize}

\end{frame}