### 2017-04-30 slides/slides.tex
[ slides ] djvu fonts added.

\documentclass[xetex, hyperref={pdfpagelabels=false}]{beamer}

% --------------------------------------------------------------------------
% Packages
% --------------------------------------------------------------------------
\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\EnableBpAbbreviations
\def\extraVskip{3pt}

\usepackage{fancyvrb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{mylst}
\usepackage{url}

% --------------------------------------------------------------------------
% Tikz Configuration
% --------------------------------------------------------------------------
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usepackage{rotating}

\tikzset{
    hyperlink node/.style={
        alias=sourcenode,
        append after command={
            let     \p1 = (sourcenode.north west),
                \p2=(sourcenode.south east),
                \n1={\x2-\x1},
                \n2={\y1-\y2} in
            node [inner sep=0pt, outer sep=0pt,anchor=north west,at=(\p1)]
            {\hyperlink{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
                    %xelatex needs \XeTeXLinkBox, won't create a link unless it
                    %finds text --- rules don't work without \XeTeXLinkBox.
                    %Still builds correctly with pdflatex and lualatex
        }
    }
}

\tikzstyle{vertex}=[auto=left,circle,fill=black!25,minimum size=5pt,inner sep=0pt]

% \usetikzlibrary{shapes.geometric,shapes.symbols,shadows}

\newcommand{\gobackreconstruction}{
  \begin{tikzpicture}[overlay, remember picture, scale=0.5]
    \tikzset{shift={(current page.center)}, xshift=11.7cm,yshift=-8.4cm}
    \node[fill=plum,text=white, hyperlink node = reconstruction]
      (go) at (0,0) {\tiny Go Back};
  \end{tikzpicture}}

% --------------------------------------------------------------------------
% Color definitions.
% --------------------------------------------------------------------------
\usepackage{xcolor}
\definecolor{aliceblue}{rgb}{0.94, 0.97, 1.0}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\definecolor{amethyst}{rgb}{0.6, 0.4, 0.8}
\definecolor{antiquefuchsia}{rgb}{0.57, 0.36, 0.51}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
\definecolor{ballblue}{rgb}{0.13, 0.67, 0.8}
\definecolor{blue(munsell)}{rgb}{0.0, 0.5, 0.69}
\definecolor{blue(pigment)}{rgb}{0.2, 0.2, 0.6}
\definecolor{blu}{RGB}{1,0,102}
\definecolor{bondiblue}{rgb}{0.0, 0.58, 0.71}
\definecolor{brightmaroon}{rgb}{0.76, 0.13, 0.28}
\definecolor{cadet}{rgb}{0.33, 0.41, 0.47}

% --------------------------------------------------------------------------
% My palette.
% --------------------------------------------------------------------------
\definecolor{energy}{RGB}{49,247,250}
\definecolor{delicate}{RGB}{67,179,223}
\definecolor{faded}{RGB}{76,117,195}
\definecolor{plum}{RGB}{87,78,164}
\definecolor{petunias}{RGB}{109,80,139}
\definecolor{letour}{RGB}{101,41,105}
% --------------------------------------------------------------------------

% --------------------------------------------------------------------------
% Beamer configuration.
% --------------------------------------------------------------------------
\usetheme{default}
\usecolortheme{default}
\usefonttheme{serif}

\beamertemplatenavigationsymbolsempty
\setbeamertemplate{navigation symbols}{}
\hypersetup{pdfpagemode=UseNone}

% footer.
\setbeamercolor{headFoot}{fg=white, bg=plum!80!black}
\setbeamertemplate{footline}{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}
    [wd=.8\paperwidth,ht=2.3ex,dp=1ex,left]{headFoot}%
    \hspace*{2ex}\textbf\insertshorttitle\hspace*{2mm}|\hspace*{2mm}\textbf\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}
    [wd=.2\paperwidth,ht=2.3ex,dp=1ex,right]{headFoot}%
    \insertframenumber{}/\inserttotalframenumber\hspace*{2ex}
  \end{beamercolorbox}}%
  \vskip 0pt%
  % \vspace*{1mm}%
}

% title in the frame
\setbeamerfont{frametitle}{size=\small,series=\bfseries}
\setbeamercolor{frametitle}{fg=white,bg=plum}

\setbeamerfont{framesubtitle}{size=\normalfont\tiny\itshape}
\setbeamercolor{framesubtitle}{fg=white, bg=plum}

\setbeamercolor{background canvas}{bg=white}
\setbeamercolor{normal text}{fg=black}

% \setbeamercolor{institute}{fg=blu}
\setbeamercolor{title}{fg=plum!80!black}
% \setbeamercolor{subtitle}{fg=blu}

% \setbeamercolor{titlelike}{fg=blu}
\setbeamerfont{footnote}{size=\tiny}
\setbeamercolor{footnote}{fg=gray}
% \setbeamercolor{block title}{bg=blue,fg=blu}
% \setbeamercolor{block body}{bg=aliceblue}
% \setbeamercolor{item}{fg=blu} % color of bullets
% \setbeamercolor{subitem}{fg=blu}
% \setbeamercolor{itemize/enumerate subbody}{fg=blu}
% \setbeamertemplate{itemize subitem}{{\textendash}}
% \setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}
% \setbeamerfont{itemize/enumerate subitem}{size=\footnotesize}

% --------------------------------------------------------------------------
% Fonts
% --------------------------------------------------------------------------

\usefonttheme{professionalfonts}
\usefonttheme{serif}
\usepackage{fontspec}
\usepackage{mathtools}
\usepackage{unicode-math}

\setmathfont[ExternalLocation=fonts/
  , BoldFont=xits-mathbold.otf
  ]{xits-math.otf}
\newfontfamily\mathfont{fonts/xits-math.otf}

\setmainfont[ExternalLocation=fonts/
  , BoldFont=SourceSansPro-Semibold.otf
  , BoldItalicFont=SourceSansPro-SemiboldIt.otf
  , ItalicFont=SourceSansPro-It.otf
  ]{SourceSansPro-Regular.otf}

\setmonofont[ExternalLocation=fonts/
  , BoldFont=SourceCodePro-Semibold.ttf
  , BoldItalicFont=SourceCodePro-SemiboldIt.ttf
  , ItalicFont=SourceCodePro-It.ttf
  ]{SourceCodePro-Regular.ttf}
\newfontfamily\SourceCodeIt{fonts/SourceCodePro-It.ttf}
\newfontfamily\SourceCode{fonts/SourceCodePro-Regular.ttf}

\newcommand{\problemtptp}[3][c]
  {\lstinputlisting[style=tptp,caption=#2, firstline=#3, label=custom#1]{#2}}
\newcommand{\solutiontstp}[2][c]
  {\lstinputlisting[style=tptp,caption=#2, firstline=5, label=custom#1]{#2}}

% --------------------------------------------------------------------------
% References
% --------------------------------------------------------------------------

\usepackage[autostyle]{csquotes}
\usepackage[
    backend=biber
  , style=authoryear-icomp
  , sortlocale=en_US
  , natbib=true
  , url=false
  , doi=false
  , isbn=false
  , eprint=false
]{biblatex}
\addbibresource{ref.bib}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}
\WarningFilter{hyperref}{Token not allowed in a PDF string}

% --------------------------------------------------------------------------
% Title and Author
% --------------------------------------------------------------------------

\title[Proof Reconstruction in Classical Propositional Logic]
  {\textbf{Proof Reconstruction in Classical Propositional Logic}
}
\subtitle{(Work in Progress)}
\date{Agda Implementors’ Meeting XXV\\
May 9-15th 2017}
\author[Jonathan Prieto-Cubides]{Jonathan Prieto-Cubides\\
Joint work with Andr\'es Sicard-Ram\'irez}
\institute{
Master in Applied Mathematics\\
Universidad EAFIT\\
Medell\'in, Colombia}
% --------------------------------------------------------------------------

\newsavebox\agdapragma

\begin{document}
\setcounter{page}{1}

\begin{frame}[plain]
\titlepage
  \begin{tikzpicture}[overlay, remember picture]
   \tikzset{shift={(current page.center)}}
    \node[xshift=-4.5cm,yshift=-3.5cm] (eafit)
      {\includegraphics[width=0.2\textwidth]{figures/eafit}};
    \node[xshift=4.5cm,yshift=-3cm] (chalmers)
      {\includegraphics[width=0.18\textwidth]{figures/chalmers}};
  \end{tikzpicture}
\end{frame}

\section{Introduction}

\begin{frame}[fragile, label=overview]{Proof Reconstruction: Overview}
\vfill
\begin{tikzpicture}
\only<1->{\node(problem) at (0,0)
 {\includegraphics[scale=0.7]{figures/problem}};
}

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{\url{http://www.gilith.com/software/metis}}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{\url{http://github.com/jonaprieto/athena}}
  };
}
\only<6->{
\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{\url{http://github.com/agda/agda}}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<9->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

\only<2->{ \draw[->, thick] (problem) to (tptp)};
\only<3->{ \draw[->, thick] (tptp) to (metis)};
\only<4->{ \draw[->, thick] (metis) to (tstp)};
\only<5->{ \draw[->, thick] (tstp) to (athena)};
\only<6->{ \draw[->, thick] (athena) to (agdafile)};
\only<7->{ \draw[->, thick] (agdafile) to (agda)};
\only<8->{ \draw[->, thick] (agda) to (verified)};
\only<9->{ \draw[->, thick, gray] (agda) to (failure)};
\end{tikzpicture}
\vfill
\end{frame}

\begin{frame}[fragile, label=reconstruction]{Proof Reconstruction: Overview}
\vfill
\begin{tikzpicture}
\node(problem) at (0,0)
   {\includegraphics[scale=0.7]{figures/problem}};
\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}};
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}};
\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}};
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}};
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};

\draw[->, thick] (problem) to (tptp);
\draw[->, thick] (tptp) to (metis);
\draw[->, thick] (metis) to (tstp);
\draw[->, thick] (tstp) to (athena);
\draw[->, thick] (athena) to (agdafile);
\draw[->, thick] (agdafile) to (agda);
\draw[->, thick] (agda) to (verified);
\draw[->, thick, gray] (agda) to (failure);
\end{tikzpicture}
\vfill
\end{frame}

\subsection{Motivation}

\begin{frame}[fragile, label=tptp-syntax]{\texttt{TPTP} Syntax}
  {Thousands of Problems for Theorem Provers}

 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0cm)
    {\includegraphics[width=0.15\textwidth]{figures/tptp}};

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF %THF, TFF,
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] version in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}

\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tptp-examples]{\texttt{TPTP} Examples}

\begin{itemize}
  \item $p ⊢ p$
\begin{tptp}
fof(a, axiom, p).
fof(goal, conjecture, p).
\end{tptp}

  \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{tptp}
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{tptp}
  \end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[label=metis]{\texttt{Metis} Theorem Prover
  \footnote{\url{http://www.gilith.com/software/metis/}}}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \href{http://www.gilith.com/software/metis/logical-kernel.txt}{\textit{6 rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs detailed proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=metis-rules]{Inference Rules of \texttt{Metis}}
\hyperlink{tstp-example}{TSTP} derivations by \texttt{Metis} exhibit these inferences
\footnote{Inference rules found in proofs of Propositional Logic theorems}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Rule} & \textbf{Purpose} \\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF\\
\texttt{clausify} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction\\
\texttt{negate} & applies negation to the formula\\
\texttt{resolve} & applies theorems of resolution\\
\texttt{simplify} & applies over a list of formula to simplify them\\
\texttt{strip} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-syntax]{\texttt{TSTP} Syntax}

 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, -0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/tstp}};

A \texttt{TSTP} derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a \hyperlink{tstp-dag}{\textbf{D}irected \textbf{A}cyclic \textbf{G}raph} where
  \begin{itemize}
    \item[\texttt{leaf}] is a formula from the \hyperlink{tptp-syntax}{\texttt{TPTP}} input
    \item[\texttt{node}] is a formula inferred from parent formula
    \item[\texttt{root}] the final derived formula
  \end{itemize}
  \item Is a list of annotated formulas with the form
  \end{itemize}

\begin{center}
  {\footnotesize
\begin{tptp}
language(name, role, formula, source [,useful info]).
\end{tptp}}
\end{center}

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents)
\end{tptp}
}
\end{center}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-example]{\texttt{TSTP} Example}

 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, -0.3cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

\begin{itemize}
  \item Proof found by \texttt{Metis} Prover for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG Example}

\vfill
By refutation, we proved $p ⊢ p$:
\begin{center}
\begin{prooftree}
\AxiomC{$p$}
\RightLabel{\texttt{\tiny assume}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\tiny strip}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\tiny negate}}
\UnaryInfC{$¬ p$}
\AxiomC{$p$}
\RightLabel{\texttt{\tiny canonicalize}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\tiny simplify}}
\BinaryInfC{$⊥$}
\RightLabel{\texttt{\tiny canonicalize}}
\UnaryInfC{$⊥$}

\end{prooftree}
% \includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=athena]{\texttt{Athena}
  \footnote{\url{http://github.com/jonaprieto/athena}}}

 \begin{tikzpicture}[overlay,remember picture]
 \tikzset{shift={(current page.north)}, xshift=0cm,yshift=0cm}
  \node[] at (0.92\textwidth, 1cm)
    {\includegraphics[width=0.15\textwidth]{figures/athena}};
\end{tikzpicture}

Is a \texttt{Haskell} program
that translates proofs given by \texttt{Metis} Prover
in \texttt{TSTP} format to \texttt{Agda} code\\

\begin{itemize}
    \item Parsing of \texttt{TSTP} language
    \item Creation and analysis of \hyperlink{tstp-dag}{\textbf{DAG}} derivations
    \item Analysis of inference rules used in the \texttt{TSTP} derivation
    \item \texttt{Agda} code generation
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Library} & \textbf{Purpose} \\
\hline
\texttt{\hyperlink{agda-prop}{\texttt{Agda-Prop}}}& axioms and theorems of Classical Propositional Logic\\
\texttt{\hyperlink{agda-metis}{\texttt{Agda-Metis}}} & versions of the inference rules used by \texttt{Metis}\\
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}

\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=agda-prop-2]{\texttt{Agda-Prop} Library
  \footnote{\url{https://github.com/jonaprieto/agda-prop}}}
\begin{itemize}
  \item Intuitionistic Propositional Logic + PEM ($Γ ⊢ ϕ ∨ ¬ ϕ$)
  \item A data type for formulas
\begin{agda}
data Prop : Set where
  Var  : Fin n → Prop           -- Variables.
  $⊤$    : Prop                   -- Top (truth).
  $⊥$    : Prop                   -- Bottom (falsum).
  _$∧$_  : ($φ$ $ψ$ : Prop) → Prop    -- Conjunction.
  _$∨$_  : ($φ$ $ψ$ : Prop) → Prop    -- Disjunction.
  _$⇒$_  : ($φ$ $ψ$ : Prop) → Prop    -- Implication.
  _$⇔$_  : ($φ$ $ψ$ : Prop) → Prop    -- Biimplication.
  $¬$_   : ($φ$ : Prop) → Prop      -- Negation.
\end{agda}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-prop]{\texttt{Agda-Prop} Library
  \footnote{\url{https://github.com/jonaprieto/agda-prop}}}
\begin{itemize}
  \item A data type for theorems
\begin{agda}
data _$⊢$_ : ($Γ$ : Ctxt)($φ$ : Prop) → Set
\end{agda}

\item Constructors
\begin{agda}
assume, axiom, weaken, $⊤$-intro, $⊥$-elim, $¬$-intro,
$¬$-elim, $∧$-intro, $∧$-proj$₁$, $∧$-proj$₂$, $∨$-intro$₁$,
$∨$-intro$₂$, $∨$-elim, $⇒$-intro, $⇒$-elim, $⇔$-intro,
$⇔$-elim$₁$, $⇔$-elim$₂$.
\end{agda}
  \item Natural deduction proofs for more than 71 theorems
\begin{agda}
$⇔$-equiv, $⇔$-assoc, $⇔$-comm, $⇒-⇔-¬∨$, $⇔-¬$-to-$¬$,
$¬⇔$-to-$¬$, $¬¬$-equiv, $⇒⇒-⇔-∧⇒$, $⇔$-trans, $∧$-assoc,
$∧$-comm, $∧$-dist, $¬∧$-to-$¬∨¬$, $¬∨¬$-to-$¬∧$, $¬∨¬-⇔-¬∧$,
subst$⊢∧₁$, subst$⊢∧₂$, $∨$-assoc, $∨$-comm, $∨$-dist,
$∨$-equiv, $¬∨$-to-$¬∧¬$, $¬∧¬$-to-$¬∨$, $∨$-dmorgan,
$¬¬∨¬¬$-to-$∨$, cnf, nnf, dnf, RAA, $\cdots$
\end{agda}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-metis]{\texttt{Agda-Metis} Library
  \footnote{\url{https://github.com/jonaprieto/agda-metis}}}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.7}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|lll|}
\hline
\textbf{Rule} & \textbf{Purpose} &\textbf{Theorem}\\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF &\texttt{atp-canonicalize}\\
\texttt{clausify} & performs clausification &\texttt{atp-clausify}\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction &\hyperlink{atp-conjunct}{\texttt{atp-conjunct}}\\
\texttt{negate} & applies negation to the formula &\texttt{atp-negate}\\
\texttt{resolve} & applies theorems of resolution &\texttt{atp-resolve}\\
\texttt{simplify} & applies over a list of formula to simplify them &\texttt{atp-simplify}\\
\texttt{strip} & splits a formula into subgoals &\texttt{atp-strip}\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:agda-metis-table}
\end{table}
\vfill
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{\texttt{Agda-Metis}: Conjunct Inference
  \footnote{\url{https://github.com/jonaprieto/agda-metis}}}
\begin{itemize}
\item Definition
  \begin{equation*}
  conjunct (ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ \phi_n, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

\item Function\footnote{\label{note-conjunct}Excerpt from the \texttt{Agda-Metis}
library available in  \texttt{ATP.Metis.Rules.Conjunct} module}:
\begin{agda}
conjunct : Prop → Prop → Prop
conjunct $(φ \wedge ψ)$ $ω$ with $⌊$ eq $φ$ $ω$ $⌋$ | $⌊$ eq ψ $ω$ $⌋$
... | true  | _     = $φ$
... | false | true  = $ψ$
... | false | false = conjunct $φ$ $ω$
conjunct $φ$ $ω$       = $φ$
\end{agda}
\item Theorem\textsuperscript{\ref{note-conjunct}}
\begin{agda}
atp-conjunct
  : ∀ $\{Γ\}$ $\{φ\}$
  → $(ω : Prop)$
  → $Γ \vdash φ$
  → $Γ \vdash$ conjunct $φ$ $ω$
\end{agda}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=atp-conjunct-proof]{A proof of \texttt{atp-conjunct} theorem}
\vfill
\begin{agda}
atp-conjunct
  : $∀ \{Γ\} \{φ\}$
  → ($ω$ : Prop)
  → $Γ ⊢ φ$
  → $Γ ⊢$ conjunct $φ\ ω$

atp-conjunct $\{Γ\}$ $\{φ ∧ ψ\}$ $ω$ $Γ⊢φ$ with $⌊$ eq $φ$ $ω$ $⌋$ | $⌊$ eq $ψ$ $ω$ $⌋$
... | true  | _     = $∧$-proj$₁$ $Γ⊢φ$
... | false | true  = $∧$-proj$₂$ $Γ⊢φ$
... | false | false =
  atp-conjunct $\{Γ = Γ\}$ $\{φ = φ\}$ $ω$ ($∧$-proj$₁$ $Γ⊢φ$)
atp-conjunct {_} {Var x}    _ = id
atp-conjunct {_} {$⊤$}      _ = id
atp-conjunct {_} {$⊥$}      _ = id
atp-conjunct {_} {$φ ∨ φ₁$} _ = id
atp-conjunct {_} {$φ ⇒ φ₁$} _ = id
atp-conjunct {_} {$φ ⇔ φ₁$} _ = id
atp-conjunct {_} {$¬\,φ$}   _ = id
\end{agda}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=agda-file]{\texttt{Agda} Code Example}
  {Generated by \texttt{Athena} Tool}

\begin{itemize}
  \item The problem is $p ∧ q ⊢ q ∧ p$
  \item In \texttt{TPTP} format
\begin{tptp}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
  \item A natural deduction proof
\begin{columns}
\begin{column}{0.8\textwidth}
 \begin{prooftree}
  \AxiomC{$ϕ ∧ ψ$}
  \RightLabel{$∧$-\texttt{proj}$₁$}
  \UnaryInfC{$ϕ$}

  \AxiomC{$ϕ ∧ ψ$}
  \RightLabel{$∧$-\texttt{proj}$₂$}
  \UnaryInfC{$ψ$}
  \RightLabel{$∧$-\texttt{intro}}
  \BinaryInfC{$ψ ∧ ϕ$}
  \end{prooftree}
\end{column}
\begin{column}{0.2\textwidth}  %%<--- here
   \begin{tikzpicture}[overlay, remember picture, scale=2]
% \tikzset{shift={(current page.north)}, xshift=cm,yshift=-2cm}
\node[vertex, fill=black] (d1) {};
\node[vertex, fill=black, right=0.5cm of d1 ] (d2) {};
\node[vertex, fill=black, below=0.2cm of d1] (d3) {};
\node[vertex, fill=black, below=0.2cm of d2] (d4) {};
\node[vertex, fill=black, below right= 0.2cm and  0.25cm of d3] (d5) {};
\draw[-, thick] (d1) to (d3);
\draw[-, thick] (d2) to (d4);
\draw[-, thick] (d3) to (d5);
\draw[-, thick] (d4) to (d5);
\end{tikzpicture}
\end{column}
\end{columns}

\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, allowframebreaks]{\texttt{TSTP} Derivation}
\begin{tptp}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
fof(subgoal_0, plain, q,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, q => p,
    inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ q,
inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, (~ q),
    inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p & q,
    inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, q,
    inference(conjunct, [], [normalize_0_1])).
fof(normalize_0_3, plain, $false,
    inference(simplify, [],
        [normalize_0_0, normalize_0_2])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_3])).
fof(negate_1_0, plain, ~ (q => p),
    inference(negate, [], [subgoal_1])).
fof(normalize_1_0, plain, ~ p & q,
    inference(canonicalize, [], [negate_1_0])).
fof(normalize_1_1, plain, p & q,
     inference(canonicalize, [], [a])).
fof(normalize_1_2, plain, p,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_3, plain, q,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_4, plain, $false,
    inference(simplify, [],
      [normalize_1_0, normalize_1_2, normalize_1_3])).
cnf(refute_1_0, plain, ($false),
    inference(canonicalize, [], [normalize_1_4])).
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=verified-example]{$p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
Definitions
\begin{agda}
p, q, a, goal, subgoal$₀$, subgoal$₁$ : Prop

-- Axiom.
a = (p $∧$ q)

-- Premise.
$Γ$ : Ctxt
$Γ$ = [ a ]

-- Conjecture.
goal = (q $∧$ p)

-- Subgoals.

subgoal$₀$ = q

subgoal$₁$ = (q $⇒$ p)
\end{agda}
\end{frame}

\begin{frame}[fragile, label=verified-example-2]{$p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
\begin{agda}
-- Axiom.
a : Prop
a = (p $∧$ q)

-- Subgoal.
subgoal$₀$ : Prop
subgoal$₀$ = q

proof$₀$ : $Γ ⊢$ subgoal$₀$
proof$₀$ =
  (RAA
    (atp-canonicalize
      (atp-simplify
        (atp-canonicalize
          (atp-strip
            (assume $\{Γ = Γ\}$ (atp-negate subgoal$₀$))))
        (atp-conjunct (q)
          (atp-canonicalize
            (weaken (atp-negate subgoal$₀$)
              (assume $\{Γ = ∅\}$ a)))))))
\end{agda}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-3]{$p ∧ q ⊢ q ∧ p$}{Reconstructed Proof}
\vfill
\begin{agda}
subgoal$₁$ : Prop
subgoal$₁$ = (q $⇒$ p)

proof$₁$ : $Γ ⊢$ subgoal$₁$
proof$₁$ =
  (RAA
    (atp-canonicalize
      (atp-simplify
        (atp-conjunct (q)
          (atp-canonicalize
            (weaken (atp-negate subgoal$₁$)
              (assume $\{Γ = ∅\}$ a))))
        (atp-simplify
          (atp-canonicalize
            (atp-strip
              (assume $\{Γ = Γ\}$ (atp-negate subgoal$₁$))))
          (atp-conjunct (p)
            (atp-canonicalize
              (weaken (atp-negate subgoal$₁$)
                (assume $\{Γ = ∅\}$ a))))))))
\end{agda}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-4]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\begin{agda}
proof : $Γ ⊢$ goal
proof =
  $⇒$-elim
    atp-splitGoal
    ($∧$-intro proof$₀$ proof$₁$)
\end{agda}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {\texttt{Metis}' Issue: \url{https://github.com/gilith/metis/issues/2}}
% {}
% \begin{tptp}
% $ cat problem.tptp
% fof(goal, conjecture,
%   ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
% \end{tptp}
\begin{tptp}
$ metis --version
metis 2.3 (release 20161108)
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=hammer]{Related Work}
\begin{block}{SledgeHammer} \citep{Paulson2007}
\begin{itemize}
\item \texttt{Isabelle}/\texttt{HOL} mature tool
\item \texttt{Metis} ported within \texttt{Isabelle}
\item Reconstruct proofs of well-known ATPs: \texttt{EProver}, \texttt{Vampire},
among others using \texttt{SystemOnTPTP} server
\end{itemize}
\end{block}

\begin{block}{Integrating \texttt{Waldmeister} into \texttt{Agda}} \citep{Foster2011}
\begin{itemize}
  \item Framework for a integration between \texttt{Agda} and ATPs
  \begin{itemize}
    \item Equational Logic
    \item Reflection Layers
  \end{itemize}
  \item Source code is not available\footnote{\url{http://simon-foster.staff.shef.ac.uk/agdaatp}}
  \end{itemize}
\end{block}
\end{frame}

\begin{lrbox}{\agdapragma}
\begin{agda}
module Or where

data _OR_ (A B : Set) : Set where
  inj1 : A → A OR B
  inj2 : B → A OR B

postulate
  A B    : Set
  OR-comm : A OR B → B OR A
{-# ATP prove OR-comm #-}
\end{agda}
\end{lrbox}

\begin{frame}[fragile]{Related Work: \texttt{Apia}}
  {Proving First-Order theorems written in \texttt{Agda} using automatic theorem provers for First-Order Logic}

\only<1>{At the moment, the communication between \texttt{Agda} and
the ATPs is unidirectional because the ATPs are being used as oracles \citep{Sicard2015}
\vfill
}

\begin{tikzpicture}
\only<1->{ \node(agdafile) at (0,0){
  \includegraphics[scale=0.8]{figures/agda-pragmas}}};
\only<1>{
  \node[fill=aliceblue, rectangle, right = 2.5cm of agdafile] (atp-pragmas){
  \scalebox{0.9}{\usebox\agdapragma}
  }
  };

\only<2->{
  \node[right=1cm of agdafile] (eagda){
  \includegraphics[scale=0.8]{figures/eagda}
  \footnote{\url{https://github.com/asr/eagda}}
  }
};

\only<2->{
\node[right=1cm of eagda] (agdai){
  \includegraphics[scale=0.8]{figures/agdai}
  }
};

\only<3->{
  \node[below=0.4cm of agdai] (apia){
  \includegraphics[scale=0.8]{figures/apia}
  \footnote{\url{https://github.com/asr/apia}}}
  };

\only<4->{
  \node[below= 0.4cm of eagda] (tptp)
    {\includegraphics[scale=0.8]{figures/tptp}}};

\only<5->{ \node[below = 0.8cm of apia] (atp)
  {\includegraphics[scale=0.8]{figures/atp}}};

\only<7->{ \node[left= 2.75cm of atp, rectangle, text=white,fill=black, align=left, font=\tiny] (proved)
{
\texttt{\$ agda Or.agda}\\
\texttt{\$ apia Or.agda --atp=online-metis}\\
\texttt{Metis---2.3 proved the conjecture}
  }
};

\only<2->{\draw[->, ultra thick] (agdafile) to (eagda)};
\only<2->{\draw[->, ultra thick] (eagda) to (agdai)};
\only<3->{\draw[->, ultra thick] (agdai) to (apia)};
\only<4->{\draw[->, very thick, dashed, gray] (apia) to (tptp)};
\only<5->{\draw[->, very thick, dashed, gray] (tptp) to (atp)};
\only<6->{\draw[->, very thick, dashed, gray] (atp) to (apia)};

\end{tikzpicture}
\end{frame}

\begin{frame}[label=pending-work]{Pending Work}
\begin{itemize}
\item There are missing cases with the \texttt{simplify} inference
\item Is not clear, how  \texttt{canonicalize} inference choose what
normal form use to transform the formulas
\item Splitting a goal in a list of subgoals is not verified yet
\end{itemize}
\end{frame}

\begin{frame}[label=future-work]{Future Work}
\begin{itemize}
  \item Integration with \texttt{Apia}
  \item Support First-Order Logic with Equality
  \item Support another prover like \texttt{EProver} or \texttt{Vampire}
\end{itemize}
\end{frame}

\begin{frame}
\end{frame}

\begin{frame}[label=references]{References}
\printbibliography
\end{frame}

\end{document}