### 2017-10-11 paper/main.tex

\documentclass[runningheads]{llncs}
\usepackage[numbers]{natbib}
% \usepackage{geometry}
% \geometry{
%   a4paper,         % or letterpaper
%   textwidth=15cm,  % llncs has 12.2cm
%   textheight=24cm, % llncs has 19.3cm
%   heightrounded,   % integer number of lines
%   hratio=1:1,      % horizontally centered
%   vratio=2:3,      % not vertically centered
% }

\usepackage{subfiles}

\setcounter{tocdepth}{5}
\usepackage{graphicx}
\usepackage{array}
\usepackage{bussproofs}
\newenvironment{bprooftree}
  {\leavevmode\hbox\bgroup}
  {\DisplayProof\egroup}
\def\extraVskip{3pt}
\def\labelSpacing{3pt}
\def\defaultHypSeparation{\hskip.2in}
\def\EnableBpAbbreviations{true}

\usepackage[title]{appendix}

\usepackage{hyphenat}
\hyphenation{proof-re-cons-truc-tion proof-ob-ject}

% -------------------------------------------------------------------
% TODO notes
% -------------------------------------------------------------------

\usepackage{xargs}
\usepackage[dvipsnames]{xcolor}
% \usepackage[pdftex,dvipsnames]{xcolor}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]
  {\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]
  {\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]
  {\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25
        ,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]
  {\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]
  {\todo[disable,#1]{#2}}

% -------------------------------------------------------------------
% Tikz Configuration
% -------------------------------------------------------------------

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usepackage{rotating}

\tikzset{
  hyperlink node/.style={
    alias=sourcenode,
    append after command={
        let     \p1 = (sourcenode.north west),
            \p2=(sourcenode.south east),
            \n1={\x2-\x1},
            \n2={\y1-\y2} in
        node [inner sep=0pt, outer sep=0pt,anchor=north west,at=(\p1)]
        {\hyperlink{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
                %xelatex needs \XeTeXLinkBox, won't create a link unless it
                %finds text --- rules don't work without \XeTeXLinkBox.
                %Still builds correctly with pdflatex and lualatex
    }
  }
}

\tikzstyle{vertex}=[
    auto=left
  , circle
  , fill=black!25
  , minimum size=5pt
  , inner sep=0pt
  ]

% -------------------------------------------------------------------
% Listings
% -------------------------------------------------------------------

\usepackage{listings}
\usepackage{lstautogobble}

\lstdefinestyle{TPTP}{
    aboveskip=10pt
  , basicstyle=\tt
  , belowskip=10pt
  , keywords=[2]{axiom
    , conjecture
    , inference
    , negated_conjecture
    , plain
    , source}
  , keywords=[3]{fof, cnf}
  , keywords=[4]{canonicalize, resolve, conjunct,
    strip, simplify, negate, clausify}
  , keywordstyle=[2]\bfseries
  , keywordstyle=[3]\bfseries
  , keywordstyle=[4]\bfseries
  , showstringspaces=false
  , keepspaces=true
}

\lstnewenvironment{tptp}
  {\lstset{style=TPTP,columns=fullflexible}\csname lst@SetFirstLabel\endcsname}
  {\csname lst@SaveFirstLabel\endcsname}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ucs}
\usepackage{autofe}
\usepackage{afterpage}
\usepackage{comment}
\usepackage{url}
\usepackage{amsmath}

\newenvironment{sketchproof}{%
  \renewcommand{\proofname}{Sketch of the Proof}\proof}{\endproof}

\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{stmaryrd} %new symbol font for tcs
\usepackage{color}
\usepackage{fancybox}

\usepackage{textcomp}
\usepackage[verbose]{newunicodechar}
\newunicodechar{∧}{\ensuremath{\wedge}}
\newunicodechar{∨}{\ensuremath{\vee}}
\newunicodechar{¬}{\ensuremath{\neg}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}}
\newunicodechar{∼}{\ensuremath{\sim}}
\newunicodechar{₀}{\ensuremath{{}_{0}}}
\newunicodechar{₁}{\ensuremath{{}_{1}}}
\newunicodechar{₂}{\ensuremath{{}_{2}}}
\newunicodechar{₃}{\ensuremath{{}_{3}}}
\newunicodechar{₄}{\ensuremath{{}_{4}}}
\newunicodechar{₅}{\ensuremath{{}_{5}}}
\newunicodechar{₆}{\ensuremath{{}_{6}}}
\newunicodechar{₇}{\ensuremath{{}_{7}}}
\newunicodechar{₈}{\ensuremath{{}_{8}}}
\newunicodechar{₉}{\ensuremath{{}_{9}}}
\newunicodechar{ᵢ}{\ensuremath{{}_{i}}}
\newunicodechar{ⱼ}{\ensuremath{{}_{j}}}
\newunicodechar{ₙ}{\ensuremath{{}_{n}}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{Γ}{\ensuremath{\Gamma}}
\newunicodechar{φ}{\ensuremath{\varphi}}
\newunicodechar{ϕ}{\ensuremath{\varphi}}
\newunicodechar{ψ}{\ensuremath{\psi}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∘}{\ensuremath{\circ}}

%\DeclareUnicodeCharacter{00A0}{~}

\usepackage{verbatim}
\usepackage{relsize}
\usepackage{fancyvrb}
% \DefineVerbatimEnvironment
%   {code}{Verbatim}
%   { fontsize=\relsize{-2}
%   , fontfamily=freemodo
%   , frame =single
%   , framesep=5mm
%   , showspaces=true
%   , obeytabs=true
%   } % {}

\setcounter{secnumdepth}{5}

\renewcommand*\rmdefault{cmr}

% -------------------------------------------------------------------
% New commands
% -------------------------------------------------------------------

\usepackage{xspace}

\newcommand{\abbre}[1]{\textsc{#1}\xspace}
\newcommand{\ATP}{\abbre{ATP}}
\newcommand{\ATPs}{\abbre{ATPs}}
\newcommand{\SAT}{\abbre{SAT}}
\newcommand{\SMT}{\abbre{SMT}}
\newcommand{\ITP}{\abbre{ITP}}
\newcommand{\CPL}{\abbre{CPL}}
\newcommand{\CNF}{\abbre{CNF}}
\newcommand{\Prop}{\abbre{Prop}}
\newcommand{\Nat}{\abbre{Nat}}

\newcommand{\name}[1]{\texttt{#1}\xspace}
\newcommand{\canonicalize}{\name{canonicalize}}
\newcommand{\clausify}{\name{clausify}}
\newcommand{\conjunct}{\name{conjunct}}
\newcommand{\negate}{\name{negate}}
\newcommand{\resolve}{\name{resolve}}
\newcommand{\simplify}{\name{simplify}}
\newcommand{\strip}{\name{strip}}
\newcommand{\tptpaxiom}{\name{axiom}}
\newcommand{\tptpconjecture}{\name{conjecture}}

\newcommand{\zero}{\name{zero}}
\newcommand{\suc}{\name{suc}}
\newcommand{\nnf}{\name{nnf}}

\newcommand{\mysplit}{split}
\newcommand{\thmsplit}{lem-split}

\newcommand{\thmresolve}{thm-resolve}

\newcommand{\prg}[1]{\texttt{#1}\xspace}
\newcommand{\Agda}{\prg{Agda}}
\newcommand{\Metis}{\prg{Metis}}
\newcommand{\Athena}{\prg{Athena}}
\newcommand{\IsabelleHOL}{\prg{IsabelleHOL}}

\newcommand{\len}[1]{\texttt{#1}\xspace}
\newcommand{\Haskell}{\len{Haskell}}
\newcommand{\TPTP}{\len{TPTP}}
\newcommand{\TSTP}{\len{TSTP}}

\newcommand{\NN}{\ensuremath{\mathbb{N}}}

% -------------------------------------------------------------------
% Rest.
% -------------------------------------------------------------------

\usepackage{url}
% \urldef{\mailsc}\path|{jprieto9, asr}@eafit.edu.co|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

% -------------------------------------------------------------------
% Hyperef.
% -------------------------------------------------------------------

\usepackage{hyperref}

\begin{document}

\mainmatter

\title{Towards Proof-Reconstruction of Problems
in Classical Propositional Logic}

\titlerunning{Proof-Reconstruction in Classical
Propositional Logic}

\author{Jonathan Prieto-Cubides%
\and Andr\'es Sicard-Ram\'irez}
%
\authorrunning{Proof-Reconstruction in Classical Propositional Logic}

\institute{Logic and Computation Research Group,\\
Universidad EAFIT, Medell\'in, Colombia\\
\email{\{jprieto9,asr\}@eafit.edu.co}}

% \toctitle{Lecture Notes in Computer Science}
% \tocauthor{Authors' Instructions}

% \listoftodos[Notes]

% TODO commands!
% \unsure
% \change
% \info
% \improvement
% \thiswillnotshow

\maketitle

\subfile{sections/abstract.tex}
\subfile{sections/introduction.tex}
\subfile{sections/preliminaries.tex}
\subfile{sections/metis-language-and-proofs.tex}
\subfile{sections/proof-reconstruction.tex}
% \subfile{sections/implementation.tex}
% \subfile{sections/related-work.tex}
% \subfile{sections/conclusions.tex}
% \subfile{sections/acknowledgements.tex}

\bibliographystyle{plainnat}
\bibliography{ref}
\addcontentsline{toc}{section}{References}

\newpage
\subfile{sections/appendices.tex}

\end{document}