### 2017-10-16 paper/sections/acknowledgements.tex
[ paper ] describing simplify.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

We thank Universidad EAFIT in Medell\'in, Colombia for funding
support. This project was a product of the Logic and Computation
Research Group at Universidad EAFIT.

We thank Joe Leslie-Hurd for his support and comments about \Metis.
We also acknowledge the work carried out by Alejandro
G\'omez-Londo\~no~\cite{Gomez-Londono2015} used as the basis
for our \TSTP parsing module.

Last but not least, we thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting us to be part of the \Agda Implementors’ Meeting XXV.
We honored to present there part of this research.

\end{document}