### 2017-10-18 paper/sections/acknowledgements.tex
[ paper ] lots of grammar fixes.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

I would like to thank to Universidad EAFIT in Medell\'in, Colombia for funding
support of this research and my master studies.
This project was a product of the Logic and Computation
Research Group at Universidad EAFIT.
My special thanks to Andrés Sicard-Ramiréz for its patience, valuable comments
during this research, and in general advices and discussions around many
topics of my interest.

I thank Joe Leslie-Hurd for his support and comments about \Metis.
I also acknowledge the work carried out by Alejandro
G\'omez-Londo\~no~\cite{Gomez-Londono2015} used as the basis
for our \TSTP parsing module.

Last but not least, we thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting us to be part of the \Agda Implementors’ Meeting XXV.
I was honored to present there part of this research.

\end{document}