### 2017-10-20 paper/sections/abstract.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\begin{abstract}
We describe a syntactical proof-reconstruction
approach to verify proofs generated by \Metis prover to problems in
classical propositional logic.
To verify these \Metis proofs, we developed a tool able to translate
such derivations to \Agda proof-terms.
We formulate in type theory each inference rule found
in \Metis derivations. This allowed us to type-check with \Agda, \Metis proofs
step-by-step. We leave it for future work,
the opportunity to extend this work to support other provers, proof assistants, and formalisms.

\keywords{proof-reconstruction, type theory, automatic theorem prover,
proof-assistant}

\end{abstract}

\end{document}