### 2017-10-22 paper/sections/acknowledgements.tex
Last changes in the Paper (#37)
* minor changes.

* [ paper ] section 2 improvements.

* [ paper ] major changes, abstract, conclusions was added.

* [ paper ] minor changes, title and abstract.

* [ paper ] fixing some mistakes.

* [ paper ] fixing error grammars and adding notes.

* [ paper ] section 3 introduction added.

* [ paper ] appendix added.

* added text.

* [ paper ] perliminaries added.

* [ paper ] implementation added, some proofs added.

* [ paper ] fixing some typos and text added.

* [ paper ] fixing some typos and text added.

* [ paper ] complexity measure added.

* [ pape

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

I would like to thank to Universidad EAFIT in Medell\'in, Colombia for funding
support of this research and my master studies.
This project was a product of the Logic and Computation
Research Group at Universidad EAFIT.
My special thanks to Andrés Sicard-Ramiréz to give me the opportunity
to work on these exciting topics but also for its patience, discussions,
and very valuable comments during all this research.

I thank Joe Leslie-Hurd for his support and comments about \Metis.
I also acknowledge the work carried out by Alejandro
G\'omez-Londo\~no~\cite{Gomez-Londono2015} used as the basis
for our \TSTP parsing module.

Last but not least, I thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting me to be part of the \Agda Implementors’ Meeting XXV.
I was honored to present there part of this research.

\end{document}