### 2017-10-22 paper/sections/proof-reconstruction/negate.tex
[ paper ] conjunct reviewed.

\documentclass[../../main.tex]{subfiles}
\begin{document}

\subsubsection{Negating subgoals.}
\label{sssec:negate}

Each proof of a subgoal is a refutation, thereby each proof assumes
the negation of its subgoal.
The \negate rule assumes the negation of such subgoals. Then, this rules
is always derived from a strip inference application as we can see in 
the following \TSTP excerpt.

\begin{verbatim}
fof(sg0, (~ p => q), inf(strip, [goal])).
fof(neg0, ~ (~ p => q), inf(negate, [sg0])).
\end{verbatim}

To emulate this rule, we only need to introduce the negation
of the subgoal by using $assume$ rule from the inferences in
Fig.~\ref{fig:CPL-inference-rules}.

\end{document}