### 2017-10-22 paper/sections/conclusions.tex
[ paper ] conclusions added.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Conclusions}
\label{sec:conclusions}

We presented a reliable proof-reconstruction in type theory
for the propositional fragment of the \Metis prover.

We built a proof-reconstruction tool written in \Haskell
to generate \Agda proof-terms of \Metis \TSTP derivations.
\Agda files delivered by the translator contain call to theorems from the
\Agda formalizations of \Metis reasoning~\cite{AgdaProp,AgdaMetis}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{ssec:emulating-inferences}.
We chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

As an early future work, we propose to extend the
proof-reconstruction presented in this paper to support the
proposition fragment with equality of \Metis.
Another option is to extend the proof-reconstruction to support other \ATPs for \CPL like \name{EProver} or \name{Z3}.
This development can be carried out by following the \name{EProver}
inference rules described on Kanso's Ph.D. Thesis~\cite{Kanso2012}.

Besides the fact, Kansos~\cite{Kanso2012} proof-reconstruction is
similar to the approach presented in Section~\ref{ssec:workflow},
the authors aim to justify some of the inference
rules by using propositional semantics. We instead perform only a syntactic analysis of the derivations as we mentioned above.

Summarizing, we provided for each \Metis inference rule a theorem in
type theory to emulate the rule reasoning. Our approach is mainly
exposed in Section~\ref{ssec:emulating-inferences}.
We detect some improvements for our approach that we leave it for future
work.
For instance, we can investigate the consequences of removing \clausify
inference by strengthen \canonicalize rule in Thm.~\ref{thm:canonicalize}
The \simplify rule is still lacking of some coverage since it is fairly complex, many cases could be omitted.
Other improvements are related with computational efficiency of
rules strongly depend on reordering tasks,
Section~\ref{sssec:reordering}.
We believe some rules like \canonicalize can get some help from an analysis done in the translator tool.

One of the main contribution of this study was
increasing the trustworthiness of the automatic prover \Metis.
Justifying a proof by a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.

\end{document}