### 2017-10-24 paper/sections/conclusions-and-future-work.tex
[ paper ] reorder conclusions with future work.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Conclusions and Future Work}
\label{sec:conclusions}

We presented a reliable proof-reconstruction in type theory
for the propositional fragment of the \Metis prover. We provided for
each \Metis inference rule a formal description in type theory.
This formalizations is mainly exposed in
Section~\ref{ssec:emulating-inferences}.

We built the \Athena translator tool written in \Haskell
that generates \Agda proof-terms of \Metis derivations.
\Agda files delivered by this translator imports
\Agda formalizations of the \Metis reasoning~\cite{AgdaProp,AgdaMetis}.
We tested successful using this tool and \Agda as we
described in Section~\ref{sec:proof-reconstruction} over a subset of
propositional problems from~\cite{Prieto-Cubides2017}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{ssec:emulating-inferences}.
Nevertheless, we chose that syntactical treatment instead of using
semantics to extend this work towards the support of first-order logic.
Recall satisfiability in first-order logic is undecidable.

Summarizing,
One of the main contribution of this study was
increasing the trustworthiness of the automatic prover \Metis.
Justifying a proof by a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task to understand the prover reasoning
can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to contribute to \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by Joe Hurd in
the \Metis 20170810 release.

\subsection*{Future work}

As an early future work, we propose to extend the
proof-reconstruction presented in this paper to support the
proposition fragment with equality of \Metis.
Another option is to extend the proof-reconstruction to support other \ATPs for \CPL like \name{EProver} or \name{Z3}.
This development can be carried out by following the \name{EProver}
inference rules described on Kanso's Ph.D. thesis~\cite{Kanso2012}.

We identify some improvements for our approach in
Section~\ref{ssec:emulating-inferences}.
For instance, we can investigate the consequences of removing \clausify
inference by strengthen \canonicalize rule in Theorem~\ref{thm:canonicalize}
The \simplify rule is still lacking of some coverage since it is fairly
complex in its implementation in the \Metis source code,
and many cases could be omitted.
Other improvements are related with computational efficiency of
rules that strongly depend on reordering tasks,
Section~\ref{sssec:resolve}.
We believe some rules like \canonicalize can get some help by
performing an analysis with \Athena.

\end{document}