### 2017-10-25 paper/sections/acknowledgements.tex
more fixes.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

Research on this paper was fully supported by Universidad EAFIT in Medellín,
Colombia. I would like to thank for this funding and its support for
my master studies.
This project was a product of the Logic and Computation
Research Group at Universidad EAFIT.
My special thanks to Andrés Sicard-Ramiréz to give me the opportunity
to work on these exciting topics but also for its patience, discussions,
and very valuable comments during all this research.

I thank Joe Leslie-Hurd for his support and comments about \Metis.
I also acknowledge the work carried out by Alejandro
G\'omez-Londo\~no~\cite{Gomez-Londono2015} used as the basis
for our \TSTP parsing module in~\cite{Athena}.

Last but not least, I thankfully acknowledge
Andreas Abel and Chalmers University of Goteborg, Sweden for
inviting me to be part of the \Agda Implementors’ Meeting XXV
where I presented part of this research in the meeting.

\end{document}