### 2017-10-26 paper/sections/proof-reconstruction/clausify.tex
[ paper ] lots of changes.

\documentclass[../../main.tex]{subfiles}
\begin{document}

\subsubsection{Clausification.}
\label{sssec:clausification}

The \clausify rule transforms a
propositional formula into its clausal normal form.
It is important to notice that this kind of conversion between one formula
to its clausal normal form is not unique, and \Metis has a customized
approach to perform that transformation. That is the reason we perform some
reordering for the \CNF of the source formula to fulfill this issue.

\begin{myexample}
In the following \TSTP derivation by \Metis, we see how
\clausify transform the \texttt{norm0} formula to get \texttt{norm1} formula.

\begin{verbatim}
fof(norm0, ¬ p ∨ (q ∧ r) ...
fof(norm1, (¬ p ∨ q) ∧ (¬ p ∨ r), inf(clausify, [norm0])).
\end{verbatim}

\end{myexample}

\begin{mainth}
\label{thm:clausify}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fclausify~φ~ψ$, where,

  \begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fclausify : \Prop → \Prop → \Prop\\
  &\begin{array}{llll}
  \fclausify~φ~ψ &=
         \begin{cases}
        ψ, &\text{ if }φ≡ψ\\
        \freorder_{∧∨}~(\fcnf~φ~ψ), &\text{ otherwise.}
      \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}
\end{mainth}

\begin{proof}
If $φ ≡ ψ$, $Γ ⊢ \fclausify~φ~ψ$ normalizes to $Γ ⊢ ψ$. The conclusion follows by applying the $\fsubst$ lemma. Otherwise, we use Lemma~\ref{lem:reorder-and-or} and Lemma~\ref{lem:cnf}.
\end{proof}

\begin{myremark}
The \clausify rule is often preceded by the \canonicalize rule.
Both rules perform \emph{Clausification} that introduces
into the problem axioms and definitions in the domain of the solution.
The clausification is widely described by
\citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}.
\end{myremark}

\end{document}