### 2017-10-27 paper/sections/related-work.tex
[ paper ] refactoring in relate-work.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Related Work}
\label{sec:related-work}

Many approaches have been proposed for proof-reconstruction and some tools have
been implemented in the last decades. We first mention some tools in type
theory and later we listed some proof-reconstruction tools for classical logic.

\citeauthor{Kanso2012} in~\cite{Kanso2012,kanso2016light} presents a proof-
reconstruction in \Agda for propositional logic. Its tool support proof-
reconstruction for \name{EProver} and \name{Z3} ATPs following a similar work-
flow as we presented in Section~\ref{ssec:workflow}. Nonetheless, his approach
employs semantics for logic equivalences. We have avoided the use of
propositions meanings towards a future work to support other logics where a
syntactical approach plays an important role~(for an example of such logics, we
refer the reader to \cite{Agudelo-Agudelo2017}).
\citeauthor{foster2011integrating}~\cite{foster2011integrating} describe  the
proof-reconstruction in \Agda for \name{Waldmeister}~\cite{hillenbrand1997}, a
prover for pure equational logic. As far as we know, no other proof-
reconstruction has been carried out in \Agda and with \Metis prover.

Another important proof-assistant in type theory is \name{Coq}~\cite{coqteam}.
we found the \name{SMTCoq}~\cite{armand2011,Ekici2017} tool which provides a
certified checker for proof witness coming from the \SMT solver
\name{veriT}~\cite{bouton2009} and adds a new tactic named verit, that calls
\name{veriT} on any \name{Coq} goal. Also for \name{Coq}, given a fixed but
arbitrary first-order signature, \citeauthor{Bezem2002} in \cite{Bezem2002}
transform a proof produced by the first-order automatic theorem prover
\name{Bliksem}~\cite{deNivelle2003} in a \name{Coq} proof term.

There are some successful attempts using proof assistants for classical logic
instead of type theory.
Let us mention some representative of such tools. This descriptions is
mainly based on~\citeauthor{Sicard-Ramirez2016}~\cite{Sicard-Ramirez2016}.

The \name{Isabelle} proof assistant has the \name{Sledgehammer} tool.
This program provides a full integration between
automatic theorem provers~\cite{blanchette2013extending,Fleury2014,bohme2010} and
\name{Isabelle/HOL}~\cite{nipkow2002isabelle}, the specialization of
\name{Isabelle} for higher-order logic.
A modular proof-reconstruction workflow is presented jointly with
the full integration of \name{Leo-II} and \name{Satallax} provers with
\name{Isabelle/HOL} in \citeauthor{Een2004}~\cite{Een2004}.

\citeauthor{Hurd1999}~\cite{Hurd1999} integrates the first-order resolution
prover \name{Gandalf} with the high-order theorem prover
\name{HOL}~\cite{norrish2017hol}.
Its \verb!GANDALF_TAC! tactic is able to reconstruct \name{Gandalf} proofs
by using a \abbre{LCF} model. For \name{HOL Light}, a version of
\name{HOL} but with a simpler logic core, the \SMT solver \name{CVC4}
was integrated. \citeauthor{kaliszyk2013}~\cite{kaliszyk2013}
reconstruct proofs from different \ATPs with the \name{PRocH} tool by
replaying detailed inference steps from the \ATPs with internal
inference methods implemented in \name{HOL Light}.

\end{document}