### 2017-10-27 paper/main.tex
Commented out the compilation date.

\documentclass[
  envcountsame
  % envcountchap
    % envcountsect    %% In order to provide the usual behaviour expected
                    %% from \newtheorem
  , runningheads
  ]{llncs}

% \setlength{\mathindent}{1cm}
% \let\remark\relax
% \let\theorem\relax
% \let\lemma\relax
% \let\definition\relax

\spnewtheorem{mainth}{Theorem}{\bfseries}{\rmfamily}
\spnewtheorem{mainlemma}[mainth]{Lemma}{\bfseries}{\rmfamily}
\spnewtheorem{mydefinition}[mainth]{Definition}{\bfseries}{\rmfamily}
\spnewtheorem{myexamplenum}[mainth]{Example}{\bfseries}{\rmfamily}

\spnewtheorem*{notation}{Notation}{\it}{\rmfamily}
\spnewtheorem*{myremark}{Remark}{\bfseries}{\rmfamily}
% \spnewtheorem*{myexample}{Example}{\bfseries}{\rmfamily}

\usepackage{geometry}   %% helps with the text overflow
\usepackage{subfiles}

\usepackage{needspace}  %% https://tex.stackexchange.com/questions/28203/
                        %% thmtools-how-to-prevent-pagebreak-between-theorem-
                        %% and-proof

\setcounter{tocdepth}{5}
\usepackage{graphicx}
\usepackage{array}

\usepackage{bussproofs}
\newenvironment{bprooftree}
  {\leavevmode\hbox\bgroup}
  {\DisplayProof\egroup}

\def\extraVskip{3pt}
\def\labelSpacing{3pt}
\def\defaultHypSeparation{\hskip.2in}
\def\EnableBpAbbreviations{true}

\usepackage[title]{appendix}

\usepackage{hyphenat}
\hyphenation{proof-re-cons-truc-tion proof-ob-ject}

% -------------------------------------------------------------------
% TODO notes
% -------------------------------------------------------------------

\usepackage{xargs}

% If we are using the tikz package we must declare the color package
% before it.
% \usepackage[svgnames,dvipsnames]{xcolor}

\usepackage[pdftex,svgnames,dvipsnames]{xcolor}

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]
  {\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]
  {\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]
  {\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25
        ,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]
  {\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]
  {\todo[disable,#1]{#2}}

% -------------------------------------------------------------------
% Tikz Configuration
% -------------------------------------------------------------------

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usepackage{rotating}

\tikzset{
  hyperlink node/.style={
    alias=sourcenode,
    append after command={
        let     \p1 = (sourcenode.north west),
            \p2=(sourcenode.south east),
            \n1={\x2-\x1},
            \n2={\y1-\y2} in
        node [inner sep=0pt, outer sep=0pt,anchor=north west,at=(\p1)]
        {\hyperlink{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
                %xelatex needs \XeTeXLinkBox, won't create a link unless it
                %finds text --- rules don't work without \XeTeXLinkBox.
                %Still builds correctly with pdflatex and lualatex
    }
  }
}

\tikzstyle{vertex}=[
    auto=left
  , circle
  , fill=black!25
  , minimum size=5pt
  , inner sep=0pt
  ]

% -------------------------------------------------------------------
% Listings
% -------------------------------------------------------------------

\usepackage{listings}
\usepackage{lstautogobble}

\lstdefinestyle{TPTP}{
    aboveskip=10pt
  , basicstyle=\tt
  , belowskip=10pt
  , keywords=[2]{axiom
    , conjecture
    , inference
    , negated_conjecture
    , plain
    , source}
  , keywords=[3]{fof, cnf}
  , keywords=[4]{canonicalize, resolve, conjunct,
    strip, simplify, negate, clausify}
  , keywordstyle=[2]\bfseries
  , keywordstyle=[3]\bfseries
  , keywordstyle=[4]\bfseries
  , showstringspaces=false
  , keepspaces=true
}

\lstnewenvironment{tptp}
  {\lstset{style=TPTP,columns=fullflexible}\csname lst@SetFirstLabel\endcsname}
  {\csname lst@SaveFirstLabel\endcsname}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{autofe}
\usepackage{afterpage}
\usepackage{comment}
\usepackage[hyphens]{url}
\usepackage{bm}

\usepackage{amssymb}
\usepackage{amsmath}

\let\proof\relax      %% https://tex.stackexchange.com/questions/43835/
\let\endproof\relax   %% conflict-between-amsthm-and-some-other-package
\usepackage{amsthm}

\newenvironment{sketchproof}{%
  \renewcommand{\proofname}{Sketch of the Proof}\proof}{\endproof}

\renewcommand\qedsymbol{$\blacksquare$}

%\usepackage{stmaryrd} %new symbol font for tcs
\usepackage{color}
\usepackage{fancybox}

% -------------------------------------------------------------------
% Unicode Characters.
% -------------------------------------------------------------------

\usepackage{textcomp}
\usepackage[verbose]{newunicodechar}

\newunicodechar{¬}{\ensuremath{\neg}\;}
\newunicodechar{Γ}{\ensuremath{\Gamma}}
\newunicodechar{γ}{\ensuremath{\gamma}}
\newunicodechar{φ}{\ensuremath{\varphi}}
\newunicodechar{ψ}{\ensuremath{\psi}}
\newunicodechar{ϕ}{\ensuremath{\varphi}}
\newunicodechar{ᵢ}{\ensuremath{{}_{i}}}
\newunicodechar{₀}{\ensuremath{{}_{0}}}
\newunicodechar{₁}{\ensuremath{{}_{1}}}
\newunicodechar{₂}{\ensuremath{{}_{2}}}
\newunicodechar{₃}{\ensuremath{{}_{3}}}
\newunicodechar{₄}{\ensuremath{{}_{4}}}
\newunicodechar{₅}{\ensuremath{{}_{5}}}
\newunicodechar{₆}{\ensuremath{{}_{6}}}
\newunicodechar{₇}{\ensuremath{{}_{7}}}
\newunicodechar{₈}{\ensuremath{{}_{8}}}
\newunicodechar{₉}{\ensuremath{{}_{9}}}
\newunicodechar{ₙ}{\ensuremath{{}_{n}}}
\newunicodechar{ℓ}{\ensuremath{\ell}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∘}{\ensuremath{\circ}}
\newunicodechar{∙}{\ensuremath{\bullet}}
\newunicodechar{∧}{\ensuremath{\wedge}}
\newunicodechar{∨}{\ensuremath{\vee}}
\newunicodechar{∼}{\ensuremath{\sim}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{⊕}{\ensuremath{\oplus}}
\newunicodechar{⊖}{\ensuremath{\ominus}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{⊻}{\ensuremath{\veebar}}
\newunicodechar{⟝}{\ensuremath{\vdash}}
\newunicodechar{⬓}{\ensuremath{\square}}
\newunicodechar{ⱼ}{\ensuremath{{}_{j}}}
\newunicodechar{λ}{\ensuremath{\lambda}}

% -------------------------------------------------------------------
% Verbatim.
% -------------------------------------------------------------------

\usepackage{verbatim}
\usepackage{relsize}
\usepackage{fancyvrb}
% \DefineVerbatimEnvironment
%   {code}{Verbatim}
%   { fontsize=\relsize{-2}
%   , fontfamily=freemodo
%   , frame =single
%   , framesep=5mm
%   , showspaces=true
%   , obeytabs=true
%   } % {}

\setcounter{secnumdepth}{5}

\renewcommand*\rmdefault{cmr}

% -------------------------------------------------------------------
% New commands
% -------------------------------------------------------------------

\usepackage{xspace}

\newcommand{\abbre}[1]{\textsc{#1}\xspace}
\newcommand{\ATPs}{\abbre{ATPs}}
\newcommand{\ATP}{\abbre{ATP}}
\newcommand{\Bool}{\abbre{Bool}}
\newcommand{\CNF}{\abbre{CNF}}
\newcommand{\CPL}{\abbre{CPL}}
\newcommand{\ITP}{\abbre{ITP}}
\newcommand{\ITPs}{\abbre{ITPs}}
\newcommand{\List}{\abbre{List}}
\newcommand{\NProp}{\abbre{Prop}}
\newcommand{\Nat}{\abbre{Nat}}
\newcommand{\Prop}{\abbre{Prop}}
\newcommand{\Source}{\abbre{Premise}}
\newcommand{\Target}{\abbre{Conclusion}}
\newcommand{\Lit}{\abbre{Lit}}
\newcommand{\SAT}{\abbre{SAT}}
\newcommand{\SMT}{\abbre{SMT}}

\newcommand{\name}[1]{\texttt{#1}\xspace}
\newcommand{\canonicalize}{\name{canonicalize}}
\newcommand{\clausify}{\name{clausify}}
\newcommand{\conjunct}{\name{conjunct}}
\newcommand{\negate}{\name{negate}}
\newcommand{\resolve}{\name{resolve}}
\newcommand{\simplify}{\name{simplify}}
\newcommand{\strip}{\name{strip}}
\newcommand{\tptpaxiom}{\name{axiom}}
\newcommand{\tptpconjecture}{\name{conjecture}}

\newcommand{\zero}{\textsf{zero}}
\newcommand{\suc}{\textsf{succ}}

\newcommand{\nnf}{\name{nnf}}

\newcommand{\prg}[1]{\texttt{#1}\xspace}
\newcommand{\Agda}{\prg{Agda}}
\newcommand{\Athena}{\prg{Athena}}
\newcommand{\IsabelleHOL}{\prg{IsabelleHOL}}
\newcommand{\Metis}{\prg{Metis}}

\newcommand{\len}[1]{\texttt{#1}\xspace}
\newcommand{\Haskell}{\len{Haskell}}
\newcommand{\TPTP}{\len{TPTP}}
\newcommand{\TSTP}{\len{TSTP}}

\newcommand{\NN}{\ensuremath{\mathbb{N}}}

\newcommand{\fun}[1]{\textsf{#1}\xspace}

\newcommand{\fassoc}{\fun{assoc}}
\newcommand{\fbuild}{\fun{build}}
\newcommand{\fcanonicalize}{\fun{canonicalize}}
\newcommand{\fcanon}{\fun{simplify}}
\newcommand{\fclausify}{\fun{clausify}}
\newcommand{\fcnf}{\fun{cnf}}
\newcommand{\fconjunctor}{\fun{disj}}
\newcommand{\fconjunct}{\fun{conjunct}}
\newcommand{\fdist}{\fun{dist}}
\newcommand{\ffactor}{\fun{factor}}
\newcommand{\fform}{\fun{form}}
\newcommand{\fmax}{\fun{max}}
\newcommand{\fndisj}{\fun{ndisj}}
\newcommand{\fnegate}{\fun{negate}}
\newcommand{\fnnfp}{\fun{nnf}₀}
\newcommand{\fnnf}{\fun{nnf}}
\newcommand{\fpolarity}{\fun{polarity}}
\newcommand{\frank}{\fun{rank}}
\newcommand{\frconj}{\fun{rconj}}
\newcommand{\frdisj}{\fun{rdisj}}
\newcommand{\freorder}{\fun{reorder}}
\newcommand{\fresolve}{\fun{resolve}}
\newcommand{\frm}{\fun{rm}}
\newcommand{\frsol}{\fun{rsol}}
\newcommand{\fsbuild}{\fun{sbuild}}
\newcommand{\fsimplify}{\fun{simplify}}
\newcommand{\fstripp}{\fun{strip}\prime}
\newcommand{\fstrip}{\fun{strip}}
\newcommand{\fsubst}{\fun{subst}}
\newcommand{\fuh}{\fun{uh}}

% latin etc. abbrev
\newcommand{\abbrev}[1]{#1} % alternative: \emph{#1}
\newcommand{\cf}{\abbrev{cf.}\ }
\newcommand{\eg}{\abbrev{e.\,g.}}
\newcommand{\Eg}{\abbrev{E.\,g.}}
\newcommand{\ie}{\abbrev{i.\,e.}}
\newcommand{\Ie}{\abbrev{I.\,e.}}
\newcommand{\etal}{\abbrev{et.\,al.}}
\newcommand{\wwlog}{w.\,l.\,o.\,g.} % \wlog is ``write into log file''
\newcommand{\Wlog}{W.\,l.\,o.\,g.}
\newcommand{\wrt}{w.\,r.\,t.}
\newcommand{\TheEnd}{\\\hspace*{0.975\textwidth}$\blacksquare$}

% -------------------------------------------------------------------
% Rest.
% -------------------------------------------------------------------

\usepackage{url}
% \urldef{\mailsc}\path|{jprieto9, asr}@eafit.edu.co|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

% -------------------------------------------------------------------
% biblatex.
% -------------------------------------------------------------------

\usepackage[%
backend=biber,
backref=true,%
% bibstyle=authoryear,%
% citestyle=authoryear-comp,%
% giveninits=false,%%
isbn=false,%
maxbibnames=6,%
% maxcitenames=4,%
]{biblatex}
\usepackage{csquotes}

\addbibresource{ref.bib}

% Remove the quotes from the title.
\DeclareFieldFormat
  [article,%
  incollection,%
  inproceedings,%
  thesis,% works for mastersthesis and phdthesis
  unpublished,%
  ]{title}{#1}

% Remove "In:" before journal names.
%
% From
% http://tex.stackexchange.com/questions/10682/suppress-in-biblatex
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}

% Remove italics from journal names.
\DeclareFieldFormat{journaltitle}{#1}

% Remove italics from proceedings names in articles in proceedings.
\DeclareFieldFormat%
  [incollection,%
  inproceedings,%
  ]{booktitle}{#1}

% Remove italics from the title names.
\DeclareFieldFormat%
 [book,%
 misc,%
 online,%
 report,%
 ]{title}{#1}

% Remove URL if there is DOI.
%
% From:
% https://tex.stackexchange.com/questions/154864/biblatex-use-doi-only-if-there-is-no-url

\renewbibmacro*{doi+eprint+url}{%
  \iftoggle{bbx:url}
    {\iffieldundef{doi}{\usebibmacro{url+urldate}}{}}
    {}%
  \newunit\newblock
  \iftoggle{bbx:eprint}
    {\usebibmacro{eprint}}
    {}%
  \newunit\newblock
    \iftoggle{bbx:doi}
    {\printfield{doi}}
    {}}

% -------------------------------------------------------------------
% Hyperef.
% -------------------------------------------------------------------

\usepackage{hyperref}

\hypersetup{citecolor=DarkGreen}

% Colors the text instead of their the anchors.
\hypersetup{colorlinks=true}

\hypersetup{urlcolor=blue}

% -------------------------------------------------------------------
% Include a footer with date and hour of the compilation
% -------------------------------------------------------------------

% \usepackage{fancyhdr}
% \usepackage[us,12hr]{datetime} % `us' makes \today behave as usual in TeX/LaTeX
% \fancypagestyle{plain}{
% \fancyhf{}
% \rfoot{Compiled on {\ddmmyyyydate\today} at \currenttime}
% \lfoot{Page \thepage}
% \renewcommand{\headrulewidth}{0pt}}
% \pagestyle{plain}

% -------------------------------------------------------------------

\begin{document}

\mainmatter

\title{Reconstructing Propositional Proofs in Type Theory}
% \title{Proof-Reconstruction in Type Theory for Propositional Logic}
\author{Jonathan Prieto-Cubides%
%\and Andr\'es Sicard-Ram\'irez
}
%

\institute{Logic and Computation Research Group,\\
Universidad EAFIT, Medell\'in, Colombia\\
\email{jprieto9@eafit.edu.co}}

% \toctitle{Lecture Notes in Computer Science}
% \tocauthor{Authors' Instructions}

% \listoftodos[Notes]

% TODO commands!
% \unsure
% \change
% \info
% \improvement
% \thiswillnotshow

\maketitle

\subfile{sections/abstract.tex}
% \tableofcontents
\subfile{sections/introduction.tex}
\subfile{sections/type-theory.tex}
\subfile{sections/metis-language-and-proofs.tex}
\subfile{sections/proof-reconstruction.tex}
% \subfile{sections/implementation.tex}
\subfile{sections/related-work.tex}
\subfile{sections/conclusions-and-future-work.tex}
\subfile{sections/acknowledgements.tex}

\printbibliography
\addcontentsline{toc}{section}{References}

% \newpage
\subfile{sections/appendices.tex}

\end{document}