### 2017-10-27 paper/sections/abstract.tex
[ paper ] proof-assistant and proof-term.

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{abstract}
We describe a syntactical proof-reconstruction approach to verify proofs
generated by \Metis prover to problems in classical propositional logic.
To verify these \Metis proofs, we developed a tool able to translate such
derivations to \Agda proof-terms. We formalise in type theory each inference
rule found in \Metis derivations. This allowed us to type-check with \Agda,
\Metis proofs step-by-step. We leave it for future work, the opportunity to
extend this work to support other provers, proof-assistants, and formalisms.

\keywords{proof-reconstruction, type theory, automatic theorem prover,
proof-assistant}
\end{abstract}

\end{document}