### 2017-11-14 paper/sections/abstract.tex
[ paper ] minor change in the abstract.

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{abstract}

We describe a syntactical proof-reconstruction approach to verify proofs
generated by \Metis prover to problems in classical propositional logic. To
verify \Metis proofs, we developed a tool and two Agda libraries able to
translate such derivations to \Agda proof-terms. We formalise in type theory
each inference rule found in the \Metis derivations. This allowed us to type-
check with \Agda, \Metis proofs step-by-step.

% We leave it for future work, the opportunity to
% extend this work to support other provers, proof-assistants, and formalisms.

\keywords{proof-reconstruction, type theory, automatic theorem prover,
proof-assistant}
\end{abstract}

\end{document}