### 2017-11-17 paper/sections/type-theory.tex
[ paper ] some corrections added.

\documentclass[../main.tex]{subfiles}
\begin{document}

\section{Type Theory}
\label{sec:type-theory}

Type theory is a formalism for the foundation of mathematics, and it has became in a key instrument to study logic and proof theory.
It was proposed in 1902 to avoid Russell's paradox and it follows
the same philosophy of constructive mathematics where the witness of
a statement is \emph{everything}. In that direction, we could say
the main actors in type theory are the statements and their proofs.
By following the Curry-Howard correspondence (see, for example,~\cite{Wadler-2015}), types as propositions.
A proposition corresponds to a type, and a proof of that propositions
corresponds to the term of that proof (henceforth proof-term).
%\emph{types} are theorems and their \emph{terms} are their proofs.

Since type theory is a formal system, we have a syntax and a set of
derivation rules. These rules enables us to derive a kind of conclusions called
\emph{judgments}. A judgment is another way to say that a term has a certain type.

\begin{notation}
  Types are written using \abbre{small caps}.  We write the type
  judgments as $a : \abbre{a}$ to denote that the term~$a$ is of
  type~$\abbre{a}$.
\end{notation}

We produce derivation trees inside the judgments using the derivation rules.
Since type theory can be seen as typed $λ$-calculus with dependent function
types, evaluation of $λ$-terms also called \emph{normalization} of the proof-
terms, are reductions using the system inference rules.

Therefore, the
proof verification task becomes in type theory as verifying that the proof-term has the correspondent type of the theorem. We know
this process as \emph{type-checking}. This feature of type theory allow us to
verify a proof generated by an \ATP by reconstructing its proof-term to type-
check the proof. For such a purpose, we use a proof assistant based on type
theory like \Agda to delegate this task.
% However, this approach has some weak points discussed in~\cite{Bezem2002}.
Some limitations from the type theory
point of view for proof-reconstruction is described in the following sections.

% ...................................................................

% section section_name (end)
\subsection{Terminating functions}
\label{ssec:structural-recursion}

To reconstruct \Metis inference rules in type theory, we observed that some
rules or their inner functions are implemented by \emph{general recursive}
functions.

Functions defined by a general recursion can not be directly translated in type
theory since it is not a guarantee they terminate. For that reason, we follow
the technique described in~\cite{Bertot2004} to avoid termination problems by
modifying the recursive functions to be \emph{structurally recursive}. % (see
\cite{Coquand1992,Abel2002,Bove2005} for more details).

This paper adopts the notion of \emph{structurally smaller}
from~\cite{Abel2002}, that is, a recursive function is structurally
recursive if it calls itself with only structurally smaller arguments.

General recursive functions can be rewrote into structurally recursive
functions by using for instance, the \emph{bounded recursion} technique.
For a yet another methods, we refer
the reader to \cite{Coquand1992,Abel2002,Bove2005}.
The bounded technique defines a new structural recursive function based on
the general recursive function by adding an argument.
The new argument is the \emph{bound}, a natural number given by the
function complexity. In other words, the added argument will represent
the number of times the function needs to call itself to get the
expected outcome.

\begin{notation}
  We use \Prop for the type of propositions.
  We write the equality as $φ ≡ ψ$, for $φ,~ψ :
  \Prop$. For this equality, we assume the reflexivity, symmetry and
  transitivity properties.
  We say $\varphi$ and $\psi$ are Two propositions are syntactically equivalent if there is a derivation from $\varphi$ to $\psi$ and
  a derivation from $\psi$ to $\varphi$.
   The type of natural numbers is called \Nat,
  and it is defined as usual, where $\zero$ and $\suc$ are its data
  constructors. We use names and symbols for the arithmetic operations
  as usual (\eg, $+$, $-$, $*$).  We use syntax sugar for $\zero$, $\suc$,
  $\suc~\zero$, $\dots$, with the decimal representation~$0$, $1$,
  $2$, $\dots$ as well.
\end{notation}

Hence, one approach to define a structural recursive function based on a general
recursive function $\rm{f}:~\abbre{a}~\to~\abbre{b}$ is to formulate a new
function $\rm{f}^{*}~:~\abbre{a}~\to~\Nat~\to~\abbre{b}$ where all of its
recursive calls are done on smaller arguments.

\begin{myexamplenum}
\label{ex:structural-recursion}

Let us consider the following example to show the bounded recursion technique
for the \fuh function. This function is used for reconstructing
a \Metis inference rule in Section~\ref{sssec:strip-a-goal}.

\begin{equation}
\label{eq:uh-definition}
\begin{aligned}
&\hspace{.495mm}\fuh_{0} : \Prop → \Prop\\
&\begin{array}{lll}
  \fuh_{0} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{0}~((φ₁ ∧ φ₂) ⇒ φ₃)\\
  \fuh_{0} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fuh_{0}~(φ₁ ⇒ φ₂) ∧ \fuh_{0}~(φ₁ ⇒ φ₃)\\
  \fuh_{0} &φ                &= φ.
\end{array}
\end{aligned}
\end{equation}

In~\eqref{eq:uh-definition}, the first two equations
call on arguments no structurally smaller.
For instance, in the first equation,
$$\fuh_{0} (φ₁ ⇒ (φ₂ ⇒ φ₃)) = \fuh_{0}~((φ₁ ∧ φ₂) ⇒ φ₃),$$
the function $\fuh_{0}$ calls on the  $(φ₁ ∧ φ₂) ⇒ φ₃$ argument in the right-hand side but this formula is not a subformula of $φ₁ ⇒ (φ₂ ⇒ φ₃)$.
Therefore, one way to translate $\fuh_{0}$ to type
theory is to define a new function using a bounded recursion.

\begin{equation}
\label{eq:uh-structured}
\begin{aligned}
&\hspace{.495mm} \fuh_{1} : \Prop → \Nat \to \Prop\\
&\begin{array}{llll}
\fuh_{1} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &(\suc~n) &= \fuh_{1}~((φ₁ ∧ φ₂) ⇒ φ₃)~n\\
\fuh_{1} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &(\suc~n) &= \fuh_{1}~(φ₁ ⇒ φ₂)~n ∧ \fuh_{1}~(φ₁ ⇒ φ₃)~n\\
\fuh_{1} &φ &n &= φ.
\end{array}
\end{aligned}
\end{equation}

We bounded the recursion calls of the $\fuh_{0}$ function using its
\emph{complexity measure}.

The complexity measure of a function is the number of steps
the function takes to finish. Since the functions defined for formulas
are based on the formula structure, we assign a complexity measure for each
of these cases in the formula structure. We refer the reader to~\cite{VanDalen1994,Agudelo-Agudelo2017} for more details about this kind of function.

In our case, we define the $\fuh_{cm}: \Prop → \Nat$ function as
the complexity measure for the $\fuh_0$ function.

\begin{equation}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{equation}

Following the technique mentioned above, we define the
function~$\fuh$, the structural
recursive definition of the function~$\fuh_{0}$.
\begin{align*}
  \label{eq:uh}
  \begin{split}
  &\fuh : \Prop → \Prop\\
  &\fuh~φ~ = \fuh_{1}~φ~(\fuh_{cm}~φ).
  \end{split}
\end{align*}

\end{myexamplenum}

For a more complete treatment of general recursion, and how to define
well-founded and well-structural recursive functions, we refer the
reader to~\cite{Bove2005}.

\subsection{Intuitionistic logic}
\label{ssec:intuitionistic-logic}

Type theory and intuitionistic logic are accompanying theories since
they are based on the same basis of constructive mathematics. Proving
propositions in these theories demands a witness construction for the
proof. Nonetheless, classical logic does not always have constructive
proofs since some of the proofs are stated by refutation.

To reconstruct proofs generated by \Metis, we have formalized in type
theory the classical propositional logic in~\cite{AgdaProp}.  A
briefly description of it is presented in \cite{VanDalen1994}. In this
formalization, we have to assume the principle of excluded middle
(henceforth \abbre{PEM}) as an axiom since \Metis is a prover for
classical logic. Assuming \abbre{PEM}, we can justify refutation
proofs by deriving from it the \emph{reductio ad absurdum} rule
(henceforth \abbre{RAA}). The \abbrev{RAA} rule is the formulation of
the principle of proof by contradiction, that is, a derivation of a
contradiction,~$⊥$, from the hypothesis~$¬ φ$, is a derivation of~$φ$.

\begin{notation}
  The $\abbre{List}$ type is the usual inductive and parametric type
  for lists.
\end{notation}

We formalise the syntactical consequence relation of \CPL by an
inductive family $\_⊢\_$ with two indexes, a set of propositions (the
premises) and a proposition (the conclusion), that is, $Γ ⊢ φ$
represents that there is derivation with conclusion $φ : \Prop$ from
the set of premises $Γ : \List~\Prop$. We implemented
in~\cite{AgdaProp} the syntactical consequence relation in a similar
way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and
elimination rules for connectives and the PEM axiom as the valid
inference rules in Fig.~\ref{fig:CPL-inference-rules}.

\begin{figure}
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{\,}
    \RightLabel{assume}
    \UnaryInfC{$Γ , φ ⊢ φ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{weaken}
    \UnaryInfC{$Γ , ψ ⊢ φ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{\,}
    \RightLabel{$⊤$-intro}
    \UnaryInfC{$Γ ⊢ ⊤$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ ⊥$}
    \RightLabel{$⊥$-elim}
    \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ , φ ⊢ ⊥$}
    \RightLabel{$¬$-intro}
    \UnaryInfC{$Γ ⊢ ¬ φ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ ¬ φ$}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{$¬$-elim}
    \BinaryInfC{$Γ ⊢ ⊥$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \AxiomC{$Γ ⊢ ψ$}
    \RightLabel{$\wedge$-intro}
    \BinaryInfC{$Γ ⊢ φ \wedge ψ$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ \wedge ψ$}
    \RightLabel{$\wedge$-proj$_1$}
    \UnaryInfC{$Γ⊢ φ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ \wedge ψ$}
    \RightLabel{$\wedge$-proj$_2$}
    \UnaryInfC{$Γ⊢ ψ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{$\vee$-intro$_1$}
    \UnaryInfC{$Γ ⊢ φ \vee ψ$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ ψ$}
    \RightLabel{$\vee$-intro$_2$}
    \UnaryInfC{$Γ ⊢ φ \vee ψ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ , φ ⊢ γ $}
    \AxiomC{$Γ , ψ  ⊢ γ$}
    \RightLabel{$\vee$-elim}
    \BinaryInfC{$Γ , φ \vee ψ ⊢ γ$}
  \end{bprooftree}\quad
  \begin{bprooftree}
    \AxiomC{$Γ , φ ⊢ ψ$}
    \RightLabel{$\Rightarrow$-intro}
    \UnaryInfC{$Γ ⊢ φ \Rightarrow ψ$}
  \end{bprooftree}
\end{equation*}

\[
%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \Rightarrow ψ$}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{$\Rightarrow$-elim}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}\qquad
\begin{bprooftree}
\AxiomC{\,}
\RightLabel{PEM}
\UnaryInfC{$Γ ⊢ φ ∨ ¬ φ$}
\end{bprooftree}
\]

% \[
% \qquad
% \begin{bprooftree}
% \AxiomC{$Γ, ¬ φ ⊢ ⊥$}
% \RightLabel{RAA}
% \UnaryInfC{$Γ ⊢ φ$}
% \end{bprooftree}
% \]

\caption{Inference rules for propositional logic.}
\label{fig:CPL-inference-rules}
\end{figure}

\end{document}