### 2017-11-20 paper/sections/abstract.tex
[ paper ] lots of changes.

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{abstract}

We describe a syntactical proof-reconstruction approach to verify derivations
generated by \Metis prover to theorems in classical propositional logic. To
verify such derivations, we formalise in type theory each inference rule
of the \Metis reasoning. We developed a tool jointly with two \Agda libraries
to translate \Metis derivations to \Agda proof-terms.
These developments allowed us to type-check with \Agda, \Metis proofs
step-by-step.

\keywords{proof-reconstruction, type theory, automatic theorem prover,
proof-assistant}
\end{abstract}

\end{document}