### 2017-11-20 paper/sections/proof-reconstruction/clausify.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}

\subsubsection{Clausify.}
\label{sssec:clausification}

The \clausify rule is another alternative way to transform a formula
into its clausal normal form but without any simplification since
remember this conversion is taking into account by the \canonicalize
rule. It is important to notice that this kind of conversion between
one formula to its clausal normal form is not unique, and \Metis has
a customized approach to perform that transformation. Hence, we
perform a reordering of the conjunctive normal form for the source
formula.

\begin{myexamplenum}
In the following \TSTP derivation by \Metis, we see how
\clausify transforms the \texttt{n₀} formula to get \texttt{n₁} formula.

\begin{verbatim}
  fof(n₀, ¬ p ∨ (q ∧ r) ...
  fof(n₁, (¬ p ∨ q) ∧ (¬ p ∨ r), inf(clausify, n₀)).
\end{verbatim}

\end{myexamplenum}

\begin{mainth}
\label{thm:clausify}
   Let $ψ : \Target$. If $Γ ⊢ φ$ then $Γ ⊢ \fclausify~φ~ψ$, where
  \begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fclausify : \Source → \Target → \Prop\\
  &\begin{array}{llll}
  \fclausify~φ~ψ &=
         \begin{cases}
        ψ, &\text{ if }φ≡ψ;\\
        \freorder_{∧∨}~(\fcnf~φ)~ψ, &\text{ otherwise.}
      \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}
\end{mainth}

\begin{proof}
If $φ ≡ ψ$, $Γ ⊢ \fclausify~φ~ψ$ normalizes to $Γ ⊢ ψ$. The conclusion follows by applying the $\fsubst$ lemma. Otherwise, we use Lemma~\ref{lem:reorder-and-or} and Lemma~\ref{lem:cnf}.
\end{proof}

\end{document}