### 2017-12-06 paper/sections/proof-reconstruction.tex
[ paper ] formalization pattern with some fixes.

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Proof-Reconstruction}
\label{sec:proof-reconstruction}

In this section, we describe one approach to reconstruct
propositional proofs. This reconstruction is a translation
process from a source system to the target system.
In our case, the system of origin, the automatic theorem prover, is
\Metis; the target system is a proof-assistant, \Agda. We choose
\Agda, but another proof-assistant with the same support of type
theory and inductive types could be used.

\subsection{Workflow}
\label{ssec:workflow}

\input{sections/diagram.tex}
The overview of the proof-reconstruction is presented as a workflow
in Fig.~\ref{fig:proof-reconstruction-workflow}. This workflow
follow the same basis of the proposed workflow presented
in~\cite{sultana2015}.

The process begins with a \TPTP file that encodes a problem in \CPL.
We use this file as the input of the \Metis prover. If the problem is
a theorem, \Metis generates a derivation of the proof in \TSTP
format.

With the \TSTP derivation from Step~2, we process the derivation with the
\Athena translator tool.
\Athena parses the \TSTP format, analyzes the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Section~\ref{ssec:metis-proofs}).
In the \Athena analysis, some unnecessary steps that introduce redundancies
and some unused input are removed from the proof-tree.
As result, we get from \Athena an \Agda file of the proof
with names of functions and theorems from
the \Agda libraries that accompany this article:
\texttt{agda-prop} and \texttt{agda-metis}.

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation generated by \Metis is correct
module \Agda and the proposed formalizations for the
propositional logic and for the \Metis reasoning.
In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user looking at the error in the \TSTP derivation
by \Metis, in the translation by \Athena, in the \Agda formalizations
mentioned above or in the type-checker~\Agda.

In the remainder part of this section, a formal description using
type theory is presented to build definitions and theorems of
functions necessary to reconstruct \Metis inference rules.

% ...................................................................

\subsection{Reconstructing \Metis inference rules}
\label{ssec:emulating-inferences}

In this section, we reconstruct each \Metis inference rule in
Table~\ref{tab:agda-metis-table}.
We begin describing the reconstruction of the \strip inference rule
since this rule  is the only one that closely follows the \Metis
source code. For the other rules, we present a pattern of the formal
description for these rules in
Example~\ref{ex:inference-rule-pattern} to reconstruct such rules.

\begin{notation}
  The function name written in \texttt{typewriter} font refers to a
  \Metis inference rule in the \TSTP derivations. The same function
  name written using
  \textsf{sans serif} font refers to our formalized version to
  reconstruct the rule and formally implemented in~\cite{AgdaMetis}.
  To increase the readability of our formal descriptions in
  functions and theorems, we adopt
  the convention to use \Source and \Target as synonyms of the \Prop
  type. The \Source type in the reconstructed rule definition refers
  to the argument in the \TSTP derivations, \ie the input formula of
  the inference rule. A formula of \Target type refers to the
  expected result of the rule found in the \TSTP derivation.
\end{notation}

\subfile{sections/proof-reconstruction/strip.tex}

As we mention in the beginning of this section, to reconstruct
the \Metis inference rules, we use the following pattern in
the formalization.

\begin{myexamplenum}
\label{ex:inference-rule-pattern}
Let \texttt{metisRule} be a \Metis inference rule. To reconstruct
this rule, we define the function \fun{metisRule} in type theory
that follows the pattern:

\begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fun{metisRule} : \Source → \Target → \Prop\\
  &\begin{array}{ll}
  \fun{metisRule}~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if the conclusion }ψ\text{ can be}\text{ derived by applying certain inference}\\
         &\text{rules to the premise }φ;\\
      φ, &\text{otherwise;}
      \end{cases}
  \end{array}
  \end{aligned}
\end{equation*}

To justify all transformations done by the \fun{metisRule} rule, we
prove its soundness with a theorem like the following:

\begin{equation*}
  \text{If }Γ ⊢ φ \text{ then }Γ ⊢ \fun{metisRule}~φ~ψ\enspace.
\end{equation*}

\end{myexamplenum}

The remainder of this section will be devoted to present a formal
description in type theory of remaining \Metis rules presented in
Table~\ref{tab:agda-metis-table} using the pattern presented in the
example above. We follow the same order to present the rules as the
table shows.

% ------------------------------------------------------------------
% Rules that follow the pattern:

\subfile{sections/proof-reconstruction/conjunct.tex}
\subfile{sections/proof-reconstruction/resolve.tex}
\subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}

% -------------------------------------------------------------------

We have finished the formalization of every inference rule in
a \Metis derivation, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena
jointly with the \Agda formalizations of the rules mentioned above
with more than eighty representative theorems in propositional
logic. An interested reader can test the
problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

\end{document}