### 2018-01-03 pubs/itp-2018/sections/abstract.tex
ITP 2018: Initial commit.

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{abstract}

We describe a syntactical proof-reconstruction approach to verify
derivations generated by \Metis prover to theorems in classical
propositional logic. To verify such derivations, we formalize in type
theory each inference rule of the \Metis reasoning. We developed a
tool jointly with two \Agda libraries to translate \Metis derivations
to \Agda proof-terms. These developments allowed us to type-check
with \Agda, \Metis derivations step-by-step.

\keywords{proof-reconstruction, type theory, automatic theorem prover,
proof-assistant}
\end{abstract}

\end{document}