### 2018-01-03 pubs/itp-2018/sections/acknowledgements.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}
The first author would like to thank to Universidad EAFIT for funding
and supporting his master studies. He thankfully acknowledge Andreas
Abel and Chalmers University of Technology at Goteborg, Sweden for
inviting him to be part of the \Agda Implementors’ Meeting XXV where
he presented part of this research. We thank Joe Leslie-Hurd for his
support and comments about \Metis.  We also acknowledge the work
carried out by Alejandro G\'omez-Londo\~no~\cite{Gomez-Londono2015}
used as the basis for our \TSTP parsing module in~\cite{Athena}. This
research was fully supported by Universidad EAFIT (grant
\#~690-000088).

\end{document}