### 2018-01-25 pubs/itp-2018/sections/abstract.tex
[ itp-2018 ] Abstract and title submitted.

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{abstract}
  We describe a syntactical proof-reconstruction approach to verify
  derivations generated by the automatic theorem prover \Metis for
  classical propositional logic theorems. To verify these derivations,
  we use \Agda, a proof assistant for dependent type theory. We
  formalise in \Agda each inference rule of \Metis reasoning, that is,
  we transform \Metis derivations into \Agda proof-terms. In a future
  full-fledged integration of \Agda with off-the-shelf automatic
  theorem provers, proof-reconstruction shall be a necessary
  component.

  \keywords{proof-reconstruction, automatic theorem provers,
    proof-assistants, Agda, Metis}
\end{abstract}

\end{document}