### 2018-02-23 pubs/itp-2018/sections/proof-reconstruction/canonicalize.tex
[ paper ] removed many examples and extra content.

\documentclass[../../main.tex]{subfiles}
\begin{document}

\subsubsection{Canonicalize.}
\label{sssec:canonicalize}

The \canonicalize rule is an inference that transforms a formula to
its negative normal form or its conjunctive normal form depending on
the role that the formula plays in the problem.
This rule jointly with the \clausify rule perform the so-called
\emph{clausification} process mainly described
in~\cite{Sutcliffe1996}.

The \canonicalize rule also removes inside of the formula
the tautologies listed in~Fig.~\ref{fig:redundancies}.

\begin{figure}
  \[
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ⊥$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ⊤$}
      \UnaryInfC{$Γ ⟝ ⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ¬ φ$}
      \UnaryInfC{$Γ ⟝ ⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ φ$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
  \]

  \[
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ⊥$}
      \UnaryInfC{$Γ ⟝ ⊥$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ⊤$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ φ$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ¬ φ$}
      \UnaryInfC{$Γ ⟝ ⊥$}
    \end{bprooftree}
  \]
  \caption{Theorems to remove inside of a formula by the
  \canonicalize rule.}
\label{fig:redundancies}
\end{figure}

The \canonicalize rule is used by \Metis to introduce the subgoals in
their refutation proofs but also helps to simplify formulas at
intermediate steps in the derivations. As far as we know, the
\canonicalize rule implements a conjunctive normal form conversion
with simplifications of tautologies. Otherwise, when
an axiom, definition or hypothesis is needed to prove some goal, this
rule gets the negative normal form of the formula.

To reconstruct the \canonicalize rule, we adapted some ideas from the
\Metis source code. The presentation of this reconstruction is as
follows. We firstly describe functions to remove tautologies inside
of the formula. After, we present the negative normal form conversion
in~Lemma~\ref{lem:nnf} and the conjunctive normal form in
Lemma~\ref{lem:cnf}. At the end of this section, we state
Theorem~\ref{thm:canonicalize} to reconstruct the \canonicalize rule.

\begin{notation}
In a disjunction, $φ ≡ φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$, we say $ψ ∈_{∨} φ$,
if there is some $i = 1, \cdots, n$ such that $ψ ≡ φᵢ$.
Note that $ψ ∈_{∨} φ$ is another representation for the equality
$ψ ≡ \freorder_{∨}~φ~ψ$.
\end{notation}

In right-associative disjunctions we remove the tautologies
in Fig.~\ref{fig:or-redundancies} using Lemma~\ref{lem:canon-or}.
We assume the formulas to be right-associative unless otherwise
stated.

\begin{figure}
\begin{equation*}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ⊥$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ⊤$}
  \UnaryInfC{$Γ ⟝ ⊤$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ φ$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ¬~φ$}
  \UnaryInfC{$Γ ⟝ ⊤$}
\end{bprooftree}\qquad
\end{equation*}
\caption{Theorems to remove inside of a disjunction by
the \canonicalize rule.}
\label{fig:or-redundancies}
\end{figure}

\begin{mainlemma}
  \label{lem:canon-or}

  Let $φ : \Prop$ be a right-associative formula. If $Γ~⟝~φ$
  then $Γ~⟝~\fcanon_{∨}~φ$, where

\begin{equation}
\label{eq:canon-or}
\begin{aligned}
 &\hspace{.495mm}\fcanon_{∨} : \NProp \to \NProp\\
 &\begin{array}{lll}
   \fcanon_{∨} &(⊥ ∨ φ)     &= \fcanon_{∨}~φ \\
   \fcanon_{∨} &(φ ∨ ⊥)     &= \fcanon_{∨}~φ \\
   \fcanon_{∨} &(⊤ ∨ φ)     &= ⊤\\
   \fcanon_{∨} &(φ ∨ ⊤)     &= ⊤\\
   % \fcanon_{∨} &(¬ φ₁ ∨ φ₂) &=
   %      \begin{cases}
   %       \fcanon_{∨}~φ₂,        &\text{ if } ¬ φ₁ ∈_{∨} φ₂;\\
   %       ⊤,                     &\text{ if } φ₁ ∈_{∨} φ₂;\\
   %       ⊤,                     &\text{ if } \fcanon_{∨}~φ₂≡ ⊤;\\
   %       ¬ φ₁,                  &\text{ if } \fcanon_{∨}~φ₂≡ ⊥;\\
   %       ¬ φ₁ ∨ \fcanon_{∨}~φ₂, &\text{ otherwise.}
   %      \end{cases}\\
   \fcanon_{∨} &(φ₁ ∨ φ₂)   &=
    \begin{cases}
     ⊤,                     & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∨} φ₂;\\
     ⊤,                     & \text{ if } (¬ φ₁) ∈_{∨} φ₂;\\
     \fcanon_{∨}~φ₂,        & \text{ if } φ₁ ∈_{∨} φ₂;\\
     ⊤,                     & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊤;\\
     φ₁,                    & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊥;\\
     φ₁ ∨ \fcanon_{∨}~φ₂,   & \text{ otherwise.}
    \end{cases}\\
   \fcanon_{∨} &φ           &= φ.
  \end{array}
\end{aligned}
\end{equation}
\end{mainlemma}

Now, we have removed tautologies in disjunctions by applying the
$\fcanon_{∨}$ function. In a similar way, we define the $\fcanon_{∧}$
function to work with conjunctions.

\begin{notation}
In a conjunction, $φ ≡ φ₁ ∧ φ₂ ∧ \cdots ∧ φₙ$, we say
$ψ ∈_{∧} φ$, if there is some $i = 1, \cdots, n$ such that $ψ ≡ φᵢ$.
Note that $ψ ∈_{∧} φ$ is another representation of
the equality $ψ ≡ \fconjunct~φ~ψ$.
\end{notation}

In right-associative conjunctions we remove the tautologies
in Fig.~\ref{fig:and-redundancies} using Lemma~\ref{lem:canon-and}.

\begin{figure}
\begin{equation*}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ⊤$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ⊥$}
  \UnaryInfC{$Γ ⟝ ⊥$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ φ$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ¬~φ$}
  \UnaryInfC{$Γ ⟝ ⊥$}
\end{bprooftree}
\end{equation*}
\caption{Conjunction tautologies.}
\label{fig:and-redundancies}
\end{figure}

\begin{mainlemma}
  \label{lem:canon-and}

  Let $φ : \Prop$ be a right-associative formula. If $Γ~⟝~φ$
  then $Γ~⟝~\fcanon_{∧}~φ$, where

  \begin{equation}
   \label{eq:canon-and}
    \begin{aligned}
     &\hspace{.495mm}\fcanon_{∧} : \NProp \to \NProp\\
      &\begin{array}{lll}
        \fcanon_{∧} &(⊥ ∧ φ)     &= ⊥  \\
        \fcanon_{∧} &(φ ∧ ⊥)     &= ⊥  \\
        \fcanon_{∧} &(⊤ ∧ φ)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ ∧ ⊤)     &= \fcanon_{∧}~φ \\
        % \fcanon_{∧} &(¬ φ₁ ∧ φ₂) &=
        %   \begin{cases}
        %     ⊥,                     &\text{ if } φ₁ ∈_{∧} φ₂;\\
        %     \fcanon_{∧}~φ₂,        &\text{ if } ¬ φ₁ ∈_{∧} φ₂;\\
        %     ⊥,                     &\text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
        %      \fcanon_{∧}~φ₂,       &\text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
        %     ¬ φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise.}
        %   \end{cases}\\
        \fcanon_{∧} &(φ₁ ∧ φ₂) &=
          \begin{cases}
            ⊥,                   & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∧} φ₂;\\
            ⊥,                   & \text{ if } (¬ φ₁) ∈_{∧} φ₂;\\
            \fcanon_{∧}~φ₂,      & \text{ if } φ₁ ∈_{∧} φ₂;\\
            φ₁,                  & \text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
            ⊥,                   & \text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
            φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise.}
          \end{cases}\\
        \fcanon_{∧} &φ         &= φ.
       \end{array}
    \end{aligned}
    \end{equation}
\end{mainlemma}

Now, we are ready to define the negative normal form of a formula with
simplifications by applying to it the $\fnnf$ function defined in
Lemma~\ref{lem:nnf} (see more details in Appendix~\ref{app:polarity-for-propositions}).
This definition is mainly based on the \Metis
source code to normalize formulas. To define such a function in type
theory we used bounded recursion as we describe in
Section~\ref{ssec:structural-recursion}.

\begin{equation}
\label{eq:nnf}
\begin{aligned}
&\hspace{.495mm}\fnnf_{1} : \Prop \to \Nat \to \Prop\\
&\begin{array}{llll}
\fnnf_{1} &(φ₁ ∧ φ₂)     &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~φ₁~n ∧ \fnnf_{1}~φ₂~n))\\
\fnnf_{1} &(φ₁ ∨ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~φ₁~n ∨ \fnnf_{1}~φ₂~n))\\
\fnnf_{1} &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ φ₂)~n))\\
\fnnf_{1} &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ (¬ φ₂))~n))\\
\fnnf_{1} &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₁) ∧ (¬ φ₂))~n))\\
\fnnf_{1} &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₂) ∧ φ₁)~n))\\
\fnnf_{1} &(¬ (¬ φ))     &(\suc~n) &= \fnnf_{1}~φ₁~n\\
\fnnf_{1} &(¬ ⊤)         &(\suc~n) &= ⊥\\
\fnnf_{1} &(¬ ⊥)         &(\suc~n) &= ⊤\\
\fnnf_{1} &φ             &\zero    &= φ
\end{array}
\end{aligned}
\end{equation}

\begin{mainlemma}
  \label{lem:nnf}
  If $Γ ⊢ φ$ then $Γ ⟝ \fnnf~φ$, where
  \begin{align*}
   \begin{split}
     &\fnnf : \NProp \to \NProp\\
     &\fnnf~φ = \fnnf_{1}~φ~(\fnnf_{cm}~φ).
   \end{split}
  \end{align*}
  The $\fnnf_{cm}$  complexity measure function
  is defined in Appendix~\ref{app:polarity-for-propositions}.
\end{mainlemma}

To get the conjunctive normal form, we make sure the formula is a
conjunction of disjunctions. For such a purpose, we use distributive laws in
Lemma~\ref{lem:dist} to get that form after applying the $\fnnf$ function.

\begin{mainlemma}
  \label{lem:dist}
  $Γ ⟝ φ$ then $Γ ⟝ \fdist~φ$, where
  \begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fdist : \NProp \to \NProp\\
  &\begin{array}{lll}
    \fdist &(φ₁ ∧ φ₂) &= \fdist~φ₁ ∧ \fdist~φ₂\\
    \fdist &(φ₁ ∨ φ₂) &= \fdist_{∨}~(\fdist~φ₁ ∨ \fdist~φ₂)\\
    \fdist &φ         &= φ
   \end{array}\\
  \text{and}\\
  &\hspace{.495mm}\fdist_{∨} : \NProp \to \NProp\\
  &\begin{array}{lll}
    \fdist_{∨}&((φ₁ ∧ φ₂) ∨ φ₃) &= \fdist_{∨}~(φ₁ ∨ φ₂) ∧ \fdist_{∨}~(φ₂ ∨ φ₃)\\
    \fdist_{∨}&(φ₁ ∨ (φ₂ ∧ φ₃)) &= \fdist_{∨}~(φ₁ ∨ φ₂) ∧ \fdist_{∨}~(φ₁ ∨ φ₃)\\
    \fdist_{∨}&φ &= φ.
    \end{array}
   \end{aligned}
  \end{equation*}
\end{mainlemma}

We get the conjunctive normal form by applying
the $\fnnf$ function follow by the $\fdist$ function.

\begin{mainlemma}
\label{lem:cnf}
  If $Γ ⊢ φ$ then $Γ ⊢ \fcnf~φ$, where
  \begin{align*}
    \begin{split}
    &\fcnf : \Prop \to \Prop\\
    &\fcnf~φ = \fdist~(\fnnf~φ).
    \end{split}
  \end{align*}
\end{mainlemma}

\begin{proof}
  Composition of Lemma~\ref{lem:dist} and Lemma~\ref{lem:nnf}.
\end{proof}

Since all the transformations in Lemma~\ref{lem:nnf} and
Lemma~\ref{lem:dist} came from logical equivalences in propositional
logic, we state the following lemmas used in the reconstruction of the \simplify rule in Lemma~\ref{lem:reduce-literal} and in
Theorem~\ref{thm:canonicalize} for the \canonicalize rule.
\begin{mainlemma}
\label{lem:nnf-inv}
  If $Γ ⊢ \fnnf~φ$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{mainlemma}
\label{lem:cnf-inv}
  If $Γ ⊢ \fcnf~φ$ then $Γ ⊢ φ$.
\end{mainlemma}

Now, we are ready to reconstruct the \fcanonicalize rule. This inference rule
defined in \eqref{eq:canonicalize} performs normalization for a proposition.
That is, depending on the role of the formula in the problem, it converts that
formula to its negative normal form or its conjunctive normal form. In both
cases, \canonicalize simplifies the formula by removing tautologies inside of
it as we widely described above for theorems in Fig.~\ref{fig:redundancies}.

Based on the \Metis source code, when the formula plays the axiom or
definition role, the \canonicalize rule transforms the source formula
to its negative normal form. Otherwise, this rule converts the
formula to a simplified conjunctive normal form.

Since this rule mostly consists of dealing with clauses, to
reconstruct this rule, our strategy mainly consists of checking the
equality of negative normal form between the source and the target
formula. If it fails, we try to reorder the conjunctive normal form
of the source formula to match with the conjunctive normal form of
the target formula. It definition is as follows.

\begin{mainth} % (fold)
  \label{thm:canonicalize}
   Let $ψ : \Target$. If $Γ ⊢ φ$ then $Γ~⊢~\fcanonicalize~φ~ψ$, where
  \begin{equation}
  \label{eq:canonicalize}
  \begin{aligned}
  &\hspace{.495mm}\fcanonicalize : \Source \to \Target \to \Prop\\
  &\fcanonicalize~φ~ψ = \begin{cases}
        ψ, &\text{ if  } ψ ≡ φ;\\
        ψ, &\text{ if  } ψ ≡ \fnnf~φ;\\
        ψ, &\text{ if  } \fcnf~ψ≡ \freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ);\\
        φ, &\text{ otherwise. }
        \end{cases}
   \end{aligned}
  \end{equation}
\end{mainth}

\begin{proof}\hspace{3mm}
\begin{itemize}
\item[∙] Case $φ ≡ ψ$. By substitution theorem we conclude $Γ ⊢ ψ$.
\item[∙] Case $ψ ≡ \fnnf~φ$.
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{Lemma~\ref{lem:nnf}}
    \UnaryInfC{$Γ ⊢~\fnnf~φ$}
    \AxiomC{$ψ ≡ \fnnf~φ$}
    \RightLabel{\fsubst}
    \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}

\item[∙] Case $\fcnf~ψ ≡ \freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$.
  \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ$}
      \RightLabel{Lemma~\ref{lem:cnf}}
      \UnaryInfC{$Γ ⊢ \fcnf~φ$}
      \RightLabel{Lemma~\ref{lem:reorder-and-or}}
      \UnaryInfC{$Γ ⊢ \freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$}
      \AxiomC{$\fcnf~ψ~≡~\freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$}
      \RightLabel{\fsubst}
      \BinaryInfC{$Γ ⊢ \fcnf~ψ$}
      \RightLabel{Lemma~\ref{lem:cnf-inv}}
      \UnaryInfC{$Γ ⊢ ψ$}
    \end{bprooftree}
  \end{equation*}
\end{itemize}
\end{proof}

\end{document}