### 2018-02-23 pubs/itp-2018/sections/proof-reconstruction/clausify.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}

\subsubsection{Clausify.}
\label{sssec:clausification}

The \clausify rule is a rule that transforms a formula into
its clausal normal form but without performing simplifications of
tautologies. Recall, this kind of conversion was already
addressed by the \canonicalize rule. It is important to notice that
this kind of conversions between one formula to its clausal normal
form are not unique, and \Metis has customized approaches to perform
that transformations. Therefore, we perform a reordering of the
conjunctive normal form given by the $\fcnf$ function defined in
Lemma~\ref{lem:cnf} with the $\freorder_{∧∨}$ function from
Lemma~\ref{lem:reorder-and-or} to the input formula of the rule.

\begin{mainth}
\label{thm:clausify}
   Let $ψ : \Target$. If $Γ ⊢ φ$ then $Γ ⊢ \fclausify~φ~ψ$, where
  \begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fclausify : \Source → \Target → \Prop\\
  &\begin{array}{llll}
  \fclausify~φ~ψ &=
         \begin{cases}
        ψ, &\text{ if }φ≡ψ;\\
        \freorder_{∧∨}~(\fcnf~φ)~ψ, &\text{ otherwise.}
      \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}
\end{mainth}

\begin{proof}
If $φ ≡ ψ$, $Γ ⊢ \fclausify~φ~ψ$ normalizes to $Γ ⊢ ψ$. The conclusion follows by applying the $\fsubst$ lemma. Otherwise, we use Lemma~\ref{lem:reorder-and-or} and Lemma~\ref{lem:cnf}.
\end{proof}

\end{document}