### 2018-02-23 pubs/itp-2018/sections/proof-reconstruction/simplify.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}

\subsubsection{Simplify.}
\label{sssec:simplify}

The \simplify rule is an inference that performs simplification of
tautologies in a list of derivations. This rule simplifies such a list
by transversing the list while applying Lemma~\ref{lem:canon-or},
Lemma~\ref{lem:canon-and}, among others theorems presented in the following.
The goal of this rule despites simplifying the list of derivations to a
new formula (often smaller than its input formulas), consists of finding
a contradiction.

We observe based on the analysis of different cases in the \TSTP
derivations that \simplify can be modeled by a function with three
arguments: two source formulas and the target formula.

Since the main purpose of the \simplify rule is simplification of
formulas, we have defined the $\freduce$ function to help removing
the negation of a given literal $ℓ$ from a input formula.

\begin{equation}
\label{def:remove-literal}
  \begin{aligned}
  &\hspace{.495mm}\freduce : \Prop → \Lit → \Prop\hspace*{3.5cm}\\
  &\begin{array}{lll}
\freduce &(φ₁~∧~φ₂) &ℓ~= \fsimplify_{∧}~(\freduce~φ₁~ℓ~∧~\freduce~φ₂~ℓ)\\
\freduce &(φ₁~∨~φ₂) &ℓ~= \fsimplify_{∨}~(\freduce~φ₁~ℓ~∨~\freduce~φ₂~ℓ)\\[2mm]
\freduce &φ &ℓ~=\begin{cases}
  ⊥,  &\text{ if }φ\text{ is a literal and }ℓ ≡ \fnnf(¬ φ);\\
  φ,  &\text{ otherwise.}
  \end{cases}
   \end{array}
  \end{aligned}
\end{equation}

\begin{mainlemma}
\label{lem:reduce-literal}
Let $ℓ$ be a literal and $φ : \Prop$. If $Γ ⊢ φ$ and $Γ ⊢ ℓ$ then
$Γ~⊢~\freduce~φ~ℓ$.
\end{mainlemma}

\begin{proof} This proof is by induction on the structure of $φ$.\\

∙~Case $φ ≡ φ₁ ∨ φ₂$.

\begin{center}
\begin{scprooftree}{1}
\AxiomC{$Γ , φ₁ ⊢ φ₁$}
\AxiomC{$Γ ⊢ ℓ$}
\RightLabel{weaken}
\UnaryInfC{$Γ , φ₁ ⊢ ℓ$}
\RightLabel{by~ind.~hyp.}
\BinaryInfC{$Γ , φ₁ ⊢ \freduce~φ₁~ℓ$}
\RightLabel{∨-intro₁}
\UnaryInfC{$Γ , φ₁ ⊢ \freduce~φ₁~ℓ ∨ \freduce~φ₂~ℓ$}
%
\AxiomC{$Γ , φ₂ ⊢ φ₂$}
\AxiomC{$Γ ⊢ ℓ$}
\RightLabel{weaken}
\UnaryInfC{$Γ , φ₂ ⊢ ℓ$}
\RightLabel{by~ind.~hyp.}
\BinaryInfC{$Γ , φ₂ ⊢ \freduce~φ₂~ℓ$}
\RightLabel{∨-intro₂}
\UnaryInfC{$Γ , φ₂ ⊢ \freduce~φ₁~ℓ ∨ \freduce~φ₂~ℓ$}
\LeftLabel{$(\mathcal{D})$\hspace{3mm}}
\RightLabel{∨-elim}
\BinaryInfC{$Γ , φ₁ ∨ φ₂ ⊢ \freduce~φ₁~ℓ ∨ \freduce~φ₂~ℓ$}
\RightLabel{⇒-intro}
\UnaryInfC{$Γ ⊢  φ₁ ∨ φ₂ ⇒ (\freduce~φ₁~ℓ ∨ \freduce~φ₂~ℓ)$}
\end{scprooftree}
\end{center}
\medskip

\def\defaultHypSeparation{\hskip.4in}
\begin{center}
\begin{scprooftree}{1}
\AxiomC{$\mathcal{D}$}
\RightLabel{⇒-elim}
\AxiomC{$Γ ⊢ φ₁ ∨ φ₂$}
\BinaryInfC{$Γ ⊢ \freduce~φ₁~ℓ ∨ \freduce~φ₂~ℓ$}
\RightLabel{Lemma~\ref{lem:canon-or}}
\UnaryInfC{$Γ ⊢ \fsimplify_{∨}~(\freduce~φ₁~ℓ ∨ \freduce~φ₂~ℓ)$}
\end{scprooftree}
\end{center}

\def\defaultHypSeparation{\hskip.2in}

∙ Otherwise use the same hypothesis $Γ ⊢ φ$.
\end{proof}

The \fsimplify function is defined in~\eqref{eq:simplify}. If some
input formula is equal to the target formula, we derive that formula.
If some formula is $⊥$, we derive the target formula by using the
⊥-elim inference rule.
Otherwise, we proceed by cases on the structure of the formula.
For example, when the second source formula is a literal, we
use the $\freduce$ function and Lemma~\ref{lem:reduce-literal}.

\begin{equation}
\label{eq:simplify}
  \begin{aligned}
  &\hspace{.495mm}\fsimplify : \Source → \Source → \Target → \Prop \hspace*{3.5cm}\\
  &\begin{array}{l}
  \fsimplify~φ₁~φ₂~ψ~=\\
  \hspace{3mm}\begin{cases}
  ψ, &\text{ if }φᵢ ≡ ⊥\text{ for some }i = 1, 2;\\
  ψ, &\text{ if }φᵢ ≡ ψ\text{ for some }i = 1, 2;\\
  \fsimplify_{∧}~(\fsimplify~φ₁~φ₂₁~ψ)~φ₂₂~ψ,
  &\text{ if }φ₂ ≡ φ₂₁ ∧ φ₂₂;\\
   \fsimplify_{∨}~(\fsimplify~φ₁~φ₂₁~ψ~∨~\fsimplify~φ₁~φ₂₂~ψ)
  &\text{ if }φ₂ ≡ φ₂₁ ∨ φ₂₂;\\
  \freduce~φ₁~φ₂, &\text{ if }φ₂\text{ is a literal};\\
  φ₁,  &\text{ otherwise.}
  \end{cases}
  \end{array}
  \end{aligned}
\end{equation}

\begin{mainlemma}
  \label{lem:binary-simplify}
  Let $ψ : \Target$. If $Γ ⊢ φ₁$ and $Γ ⊢ φ₂$ then
  $Γ ⊢ \fsimplify~φ₁~φ₂~ψ$.
\end{mainlemma}

\begin{proof}
This proof is by induction on the structure of $φ$. \\

∙~Case $φ₂ ≡ φ₂₁ ∨ φ₂₂$.

\def\defaultHypSeparation{\hskip.2in}

\begin{center}
\begin{scprooftree}{1}
\AxiomC{$Γ ⊢ φ₁$}
\RightLabel{weaken}
\UnaryInfC{$Γ , φ₂₁ ⊢ φ₁$}
\AxiomC{$Γ , φ₂₁ ⊢ φ₂₁$}
\RightLabel{by~ind.~hyp.}
\BinaryInfC{$Γ , φ₂₁ ⊢ \fsimplify~φ₁~φ₂₁~ψ$}
\LeftLabel{$(\mathcal{D}_{1})$\hspace{5mm}}
\RightLabel{∨-intro₁}
\UnaryInfC{$Γ , φ₂₁ ⊢ \fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₁~φ₂₁~ψ$}
\end{scprooftree}
\end{center}
\medskip
\begin{center}
\begin{scprooftree}{1}
\AxiomC{$Γ ⊢ φ₁$}
\RightLabel{weaken}
\UnaryInfC{$Γ , φ₂₂ ⊢ φ₁$}
\AxiomC{$Γ , φ₂₂ ⊢ φ₂₂$}
\RightLabel{by~ind.~hyp.}
\BinaryInfC{$Γ , φ₂₂ ⊢ \fsimplify~φ₁~φ₂₂~ψ$}
\LeftLabel{$(\mathcal{D}_{2})$\hspace{5mm}}
\RightLabel{∨-intro₂}
\UnaryInfC{$Γ , φ₂₂ ⊢ \fsimplify~φ₁~φ₂₂~ψ ∨ \fsimplify~φ₁~φ₂₂~ψ$}
\end{scprooftree}
\end{center}
\medskip
\begin{center}
\begin{scprooftree}{1}
\AxiomC{$\mathcal{D}_{1}$}
% \UnaryInfC{$Γ , φ₂₁ ⊢ \fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₁~φ₂₁~ψ$}
\AxiomC{$\mathcal{D}_{2}$}
% \UnaryInfC{$Γ , φ₂₂ ⊢ \fsimplify~φ₁~φ₂₂~ψ ∨ \fsimplify~φ₁~φ₂₂~ψ$}
\RightLabel{∨-elim}
\BinaryInfC{$Γ , φ₂₁ ∨ φ₂₂ ⊢ \fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₁~φ₂₂~ψ$}
\RightLabel{⇒-intro}
\LeftLabel{$(\mathcal{D}_{3})$\hspace{5mm}}
\UnaryInfC{$Γ ⊢  φ₂₁ ∨ φ₂₂ ⇒ (\fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₁~φ₂₂~ψ)$}
\end{scprooftree}
\end{center}

\medskip
\begin{center}
\begin{scprooftree}{1}
\AxiomC{$\mathcal{D}_{3}$}
\UnaryInfC{$Γ ⊢  φ₂₁ ∨ φ₂₂ ⇒ (\fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₁~φ₂₂~ψ)$}
\RightLabel{⇒-elim}
\AxiomC{$Γ ⊢ φ₂₁ ∨ φ₂₂$}
\BinaryInfC{$Γ ⊢ \fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₂~φ₂₂~ψ$}
\RightLabel{Lemma~\ref{lem:canon-or}}
\UnaryInfC{$Γ ⊢ \fsimplify_{∨}~(\fsimplify~φ₁~φ₂₁~ψ ∨ \fsimplify~φ₂~φ₂₂~ψ)$}
\end{scprooftree}
\end{center}
\medskip

∙~Case $φ₂$ is a literal. Use Lemma~\ref{lem:reduce-literal} with $φ = φ₁$ and
$ℓ = φ₂$.
\end{proof}

\begin{mainth}
  \label{thm:simplify}
  Let $ψ : \Target$. Let $φᵢ : \Source$ such that $Γ ⊢ φᵢ$ for
  $i = 1, \cdots, n$ and $n \geq 2$.
  Then $Γ ⊢ \fsimplify~γ_{n-1}~φ_{n}~ψ$ where
  $γ₁ = φ₁$, and $γᵢ ≡ \fsimplify~γ_{i-1}~φᵢ~ψ$.
\end{mainth}

\begin{proof} We prove this theorem by induction on $n$.\\
∙~Case $n = 2$. Use Lemma~\ref{lem:binary-simplify}.\\
∙~Case $n > 2$. Suppose this theorem is valid for $n$, that is,
for $i = 1, \cdots, n$, $Γ ⊢ \fsimplify~γ_{n-1}~φ_{n}~ψ$. Let us prove it for $n+1$.\\

\begin{equation*}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fsimplify~γ_{n-1}~φ_{n}~ψ$}
\AxiomC{$γ_{n} ≡ \fsimplify~γ_{n-1}~φ_{n}~ψ$}
\RightLabel{subst}
\BinaryInfC{$Γ ⊢ γ_{n}$}
\AxiomC{$Γ ⊢ φ_{n+1}$}
\RightLabel{by~ind.~hyp.}
\BinaryInfC{$Γ ⊢ \fsimplify~γ_{n}~φ_{n+1}~ψ$}
\end{bprooftree}
\end{equation*}
\end{proof}

\begin{myremark}

Besides the fact that $\List~\Prop~\to~\Prop$ is the type that mostly
fit with the \simplify rule, we choose a different option. In the
translation from \TSTP to \Agda, we take the list of derivations and
we apply the rule by using a left folding (the \verb!foldl! function
in functional programming) with the $\fsimplify$ function over the
list of $φ₁, φ₂, \cdots, φₙ$. This design decision avoids us to define a new theorem
type to support \List~\Prop type in the premise side.
\end{myremark}

\begin{myexamplenum}
Let us review the following \TSTP excerpt where \simplify was used twice.

\begin{verbatim}
  fof(n₀, (¬ p ∨ q) ∧ ¬ r ∧ ¬ q ∧ (p ∨ (¬ s ∨ r)), ...
  fof(n₁, p ∨ (¬ s ∨ r), inf(conjunct, n₀)).
  fof(n₂, ¬ p ∨ q, inf(conjunct, n₀)).
  fof(n₃, ¬ q, inf(conjunct, n₀)).
  fof(n₄, ¬ p, inf(simplify, [n₂, n₃])).
  fof(n₅, ¬ r, inf(conjunct, n₀)).
  fof(n₆, ⊥, inf(simplify, [n₁, n₄, n₅])).
\end{verbatim}

\begin{enumerate}
\item The \simplify rule derives $¬ p$ in \verb!n₄!
from \verb!n₂! and \verb!n₃! derivations.

$$\fsimplify~(¬ p ∨ q)~(¬ q)~(¬ p) = ¬ p.$$
% \begin{verbatim}
% fof(norm2, ¬ p ∨ q, inf(conjunct, [norm0])).
% fof(norm3, ¬ q, inf(conjunct, [norm0])).
% fof(norm4, ¬ p, inf(simplify, [norm2, norm3])).
% \end{verbatim}
\item We use Theorem~\ref{thm:simplify} to derive ⊥ in \verb!n₆!, the
following is a proof.

\begin{equation*}
\begin{bprooftree}
\AxiomC{$Γ ⊢ p ∨ (¬ s ∧ r)$}
\AxiomC{$Γ ⊢ ¬ p$}
\RightLabel{Theorem~\ref{thm:simplify}}
\BinaryInfC{$Γ ⊢ ¬ s ∧ r$}
\AxiomC{$Γ ⊢ ¬ r$}
\RightLabel{Theorem~\ref{thm:simplify}}
\BinaryInfC{$Γ ⊢ ⊥$}
\end{bprooftree}
\end{equation*}

% \begin{verbatim}
% fof(n1, g ∨ (¬ s ∧ r), inf(conjunct, [n0])).
% fof(n4, ¬ p, inf(simplify, [n2, n3])).
% fof(n5, ¬ r, inf(conjunct, [n0])).
% fof(n6, ⊥, inf(simplify, [n1, n4, n5])).
% \end{verbatim}
\end{enumerate}
\end{myexamplenum}

% subsubsection simplify (end)
% -------------------------------------------------------------------

\end{document}