### 2018-03-13 pubs/itp-2018/sections/metis-language-and-proofs.tex
Complete first revision to section about Metis

%
\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

% \begin{myremark}
% We address proof-reconstruction in this paper for refutation proofs.
% We annotate \abbre{CPL} problems using \verb!fof! formulas that
% \Metis converts to \abbre{CNF} clauses to look for a refutation.
% \end{myremark}

\section{Metis Prover}
\label{sec:metis-language-and-proofs}

\Metis is an automatic theorem prover for first-order logic with
equality developed by Hurd~\cite{hurd2003first}. This prover is
suitable for proof-reconstruction since it provides well-documented
proofs to justify its deduction steps from the basis of only six
inference rules for first-order logic (see, for example,
\cite{paulson2007source,Farber2015}). For the propositional fragment,
\Metis has three inference rules, see Fig.~\ref{fig:metis-inferences}.

\begin{figure}
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{}
    \RightLabel{axiom~$φ₁,\cdots,φₙ$}
    \UnaryInfC{$Γ ⊢ φ₁ ∨ \cdots ∨ φₙ$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{}
    \RightLabel{assume $φ$}
    \UnaryInfC{$Γ ⊢ φ ∨ ¬ φ$}
  \end{bprooftree}
  \end{equation*}
  \vskip2mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ ℓ ∨ φ$}
    \AxiomC{$Γ ⊢ ¬ ℓ ∨ ψ$}
    \RightLabel{resolve $ℓ$}
    \BinaryInfC{$Γ ⊢ φ ∨ ψ$}
  \end{bprooftree}
\end{equation*}
\caption{\Metis inference rules for propositional logic.}
\label{fig:metis-inferences}
\end{figure}

% -------------------------------------------------------------------

\subsection{Input language}
\label{ssec:input-language}

The \TPTP language is the input language to encode problems used by \Metis.
It includes the first-order form (denoted by \name{fof}) and clause normal
form (denoted by \name{cnf}) formats~\cite{sutcliffe2009}.
The \TPTP syntax\footnote{See the complete syntax grammar
at \url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}} is a
well-defined grammar to handle annotated formulas with the form:
\begin{verbatim}
  language(name, role, formula).
\end{verbatim}
where the \name{language} can be \name{fof} or \name{cnf} and the \name{name} serves
to identify the formula within the problem. Each
annotated formula also assumes a \name{role}. Some roles frequently used are \name{axiom},
\name{conjecture}, \name{definition}, \name{plain} and \name{hypothesis} but the reader
can find the full list in~\cite{sutcliffe2009}.

On the other hand, the well-written formulas can include 
the constants \verb!$true! and \verb!$false!, the negation unary
operator~(\verb!~!), and the binary connectives
(\verb!&!, \verb!|!, \verb!=>!) to represent
($⊤$, $⊥$, $¬$, $∧$, $∨$, $⇒$) respectively.
Finally, to refer us to propositional variables,
we can just write a word using letters and numbers.
%http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html#atomic_word

\subsection{Output language}
\label{ssec:output-language}

The \TSTP language is an output language for derivations used by 
\Metis and other ATPs~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}.
A \TSTP  derivation generated by \Metis encodes a natural
deduction proof using only as valid deduction steps the inference rules
in Fig.~\ref{tab:agda-metis-table}. 
These derivations are directed acyclic graphs, trees of refutations.
Each node stands for an application of an inference rule to the formulas
in the parent nodes. The leaves in the tree represent formulas in the given
\TPTP input. Each node is labeled with a name of its inference rule (\eg, \canonicalize).
Each edge in the tree links a premise with one conclusion.

Since \Metis attempts to prove conjectures by refutation, and the root
is the final derived formula in these proof trees,  we have at their
roots the conclusion $⊥$. Such derivations are a list of annotated
formulas with the form:

\begin{verbatim}
  language(name, role, formula, source [,useful info]).
\end{verbatim}

The new field regarding the \TPTP syntax is the \name{source} field
that mainly intends to record the inference rule of the deduction step.
This field has the following pattern:

\begin{verbatim}
  inference(rule, useful info, parents).
\end{verbatim}

We have the \name{rule} field with the inference name, another field
including supporting arguments to apply the reasoning step, and the
last one field with a list its parents nodes.

For increasing the readability of this paper, we are not presenting
these \TSTP derivations using the above description. Instead of it, we
have adopted a customized \TSTP syntax to keep as short as possible
the \Metis derivations. For instance,  we shortened some keywords,
removed unnecessary arguments, and replaced some symbols. The list of
all customizations is in Appendix~\ref{app:tstp-syntax}.

\begin{myexamplenum}
\label{fig:metis-example-tree}

Let us consider the following \TSTP derivation generated by \Metis after using
the customized \TSTP syntax proposed for the theorem:

\begin{equation}
(p \Rightarrow q) \wedge (q \Rightarrow p) ⊢ (p \vee q) \Rightarrow (p \wedge q).
\end{equation}

% this tstp example should be used to explain the reconstruction.
\begin{verbatim}
  fof(premise, axiom, (p ⇒ q) ∧ (q ⇒ p)).
  fof(goal, conjecture, (p ∨ q) ⇒ (p ∧ q)).
  fof(s₀, (p ∨ q) ⇒ p, inf(strip, goal)).
  fof(s₁, ((p ∨ q) ∧ p) ⇒ q, inf(strip, goal)).
  fof(neg₀, ¬ ((p ∨ q) ⇒ p), inf(negate, s₀)).
  fof(n₀₀, (¬ p ∨ q) ∧ (¬ q ∨ p), inf(canonicalize, premise)).
  fof(n₀₁, ¬ q ∨ p, inf(conjunct, n₀₀)).
  fof(n₀₂, ¬ p ∧ (p ∨ q), inf(canonicalize, neg₀)).
  fof(n₀₃, p ∨ q, inf(conjunct, n₀₂)).
  fof(n₀₄, ¬ p, inf(conjunct, n₀₂)).
  fof(n₀₅, q, inf(simplify, [n₀₃, n₀₄])).
  cnf(r₀₀, ¬ q ∨ p, inf(canonicalize, n₀₁)).
  cnf(r₀₁, q, inf(canonicalize, n₀₅)).
  cnf(r₀₂, p, inf(resolve, q, [r₀₁, r₀₀])).
  cnf(r₀₃, ¬ p, inf(canonicalize, n₀₄)).
  cnf(r₀₄, ⊥, inf(resolve, p, [r₀₂, r₀₃])).
  fof(neg₁, ¬ ((p ∨ q) ∧ p) ⇒ q), inf(negate, s₁)).
  fof(n₁₀, ¬ q ∧ p ∧ (p ∨ q), inf(canonicalize, neg₁)).
  fof(n₁₁, (¬ p ∨ q) ∧ (¬ q ∨ p), inf(canonicalize, premise)).
  fof(n₁₂, ¬ p ∨ q, inf(conjunct, n₁₁)).
  fof(n₁₃, ⊥, inf(simplify,[n₁₀, n₁₂])).
  cnf(r₁₀, ⊥, inf(canonicalize, n₁₃)).
\end{verbatim}

\end{myexamplenum}

% -------------------------------------------------------------------

\subsection{Inference rules}
\label{ssec:metis-inferences-rules}

We present the list of inference rules used by \Metis in the \TSTP derivations
for propositional logic in Table~\ref{tab:agda-metis-table}. We reconstruct
these rules in Section~\ref{sec:proof-reconstruction}. The reader may notice
that the inference rules presented in Fig~\ref{fig:metis-inferences} diverge
from the rules in aforementioned table. The former rules are implemented by the
latter rules in the \TSTP derivations.
For instance, as far as we know, in \TSTP
derivations,  the \emph{axiom} rule is implemented by the rules
\canonicalize, \clausify, \conjunct, and \simplify.

We first present the reconstruction of the \strip inference rule and its
description since it is the rule that appears
first after each conjecture in \TSTP derivations. The other rules are sorted mainly follow their
level of complexity of their definitions and the formalization presented in
Section~\ref{ssec:emulating-inferences}. Some inference rules depend on the
formalizations of other rules. For instance, the \simplify rule and the
\clausify rule need theorems developed for the \canonicalize rule. The
\canonicalize rule needs theorems developed in the \resolve formalization,
and \resolve depends on the \conjunct rule.

% \Metis looks for a refutation to prove \verb!fof! problems.
% A \verb!fof! problem contains the conjecture, the \emph{goal}, in the
% problem. This goal is stripped into subgoals by using the \strip rule. The
% subgoals are the conjuncts in a general conjunction that is equivalent to the
% aforementioned goal.
% Therefore, to prove the goal, \Metis proves independently each subgoal since
% they tend to be less complex or even fewer propositions. At the end, the
% conjunction of all subgoal proofs must prove the goal. Nevertheless, this last
% reasoning is not explicit neither in the \TSTP derivations nor in the
% documentation available about \Metis.
% Since \Metis replies refutation proofs, to prove each subgoal, in the \TSTP
% derivation, the inference rule \negate introduces the negation of such
% subgoals. The \negate rule is a specification of the \emph{assume} rule.

\begin{table}[!ht]

\caption{\Metis inference rules.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \label{tab:agda-metis-table}
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{4mm}}l@{\hspace{2mm}}c@{\hspace{2mm}}|}
    \hline
    \textbf{Rule} & \textbf{Purpose} &\textbf{Theorem number}\\ \hline

      \texttt{strip}
      &Strip a goal into subgoals
      &\ref{thm:strip}
      \\

      \texttt{conjunct}
      &Takes a formula from a conjunction
      &\ref{thm:conjunct}
      \\

      \texttt{resolve}
      &A general form of the resolution theorem
      &\ref{thm:resolve}
      \\

      \texttt{canonicalize}
      &Normalization of the formula
      &\ref{thm:canonicalize}
      \\

      \texttt{clausify}
      &Performs clausification
      &\ref{thm:clausify}
      \\

      \texttt{simplify}
      &Simplify definitions and theorems
      &\ref{thm:simplify}
      \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

\end{document}