### 2018-03-27 pubs/itp-2018/sections/proof-reconstruction.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Proof-Reconstruction}
\label{sec:proof-reconstruction}

In this section, we describe our approach to reconstruct proofs from
\CPL derivations generated by \Metis. This reconstruction consists of
a translation from a source system to the target system. The system of
origin is the automatic theorem prover \Metis and the target system is
the proof-assistant \Agda. We choose \Agda but another proof-assistant
with the same support of type theory and inductive types could be used
(\eg \Coq).

\subsection{Workflow}
\label{ssec:workflow}

\input{sections/diagram.tex}
The overview of the proof-reconstruction is presented as the workflow
in Fig.~\ref{fig:proof-reconstruction-workflow} following the same
basis of the proposed workflow presented in~\cite{sultana2015}.

The process begins with a \TPTP file that encodes a \CPL conjecture.
Then, we use the \Metis prover to generate a derivation of the proof
in \TSTP format in case the \CPL conjecture corresponds to a theorem.
If so, the process continues. The next step is in charge of the
\Athena translator tool that generates an \Agda file.
Athena first parses the \TSTP file, then analyzes its
derivation and generates a representation of the natural deduction
proof. In the \Athena analysis, some unnecessary steps are removed
mainly because of they introduce redundancies or they do not appear in
the proof-tree. The Agda file genereted by \Athena uses names of functions
and theorems from the \Agda libraries that accompany this article:
\texttt{agda-prop} and \texttt{agda-metis}.

Finally, we type-check the proof-terms of the \Agda file.
If this type-checking success, the \TSTP derivation generated by \Metis will be
\emph{correct} module the \Agda correctness and the completeness of the
proposed formalisation and \Agda will output an interface file.

Otherwise, when the type-checking fails, the failure must be
investigated by the user. The possible causes can be due to
a \emph{completeness} issue in our \Agda libraries for the \Metis reasoning
but also in a translation bug in \Athena. We have also consider
errors due to some failure in the type-checker~\Agda,
soundness bugs in the \Metis prover.

In the remainder part of this section, we present a formal description
of definitions and theorems necessary to reconstruct \Metis inference
rules. Through this description, we refer to the lines in the \Metis
derivation presented in Fig.~\ref{complete-metis-example} of the
theorem in \eqref{eq:complete-example-problem} to show each rule.

% ...................................................................

\subsection{Reconstructing \Metis inference rules}
\label{ssec:emulating-inferences}

In this section, we reconstruct each \Metis inference rule in
Table~\ref{tab:agda-metis-table}.
We begin describing the reconstruction of the \strip inference rule
since this rule  is the only one that closely follows the \Metis
source code. To reconstruct the other rules, we use a pattern for their
formalization showed in Example~\ref{ex:inference-rule-pattern}.

\begin{notation}
  The function name written in \texttt{typewriter} font refers to a
  \Metis inference rule in the \TSTP derivations. The same function
  name written using
  \textsf{sans serif} font refers to our formalized version to
  reconstruct the rule and formally implemented in~\cite{AgdaMetis}.
  To increase the readability for functions and theorems, we use
  the convention \Source and \Target as synonyms of the \Prop
  type. The \Source type in the reconstructed rule definition refers
  to the argument in the \TSTP derivation, \ie the input formula of
  the inference rule. A formula of \Target type refers to the
  expected result of the rule found in the \TSTP derivation.
\end{notation}

\subfile{sections/proof-reconstruction/strip.tex}

Now, as we mention in the beginning of this section, to reconstruct
the \Metis inference rules, we use the following pattern in
the formalization.

\begin{myexamplenum}
\label{ex:inference-rule-pattern}
Let \texttt{metisRule} be a \Metis inference rule. To reconstruct
this rule, we define the function \fun{metisRule} in type theory
that follows the pattern:

\begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fun{metisRule} : \Source → \Target → \Prop\\
  &\begin{array}{ll}
  \fun{metisRule}~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if the conclusion }ψ\text{ can be}\text{ derived by applying certain inference}\\
         &\text{rules to the premise }φ;\\
      φ, &\text{otherwise;}
      \end{cases}
  \end{array}
  \end{aligned}
\end{equation*}

To justify all transformations done by the \fun{metisRule} rule, we
prove its soundness with a theorem like the following:

\begin{equation*}
  \text{If }Γ ⊢ φ \text{ then }Γ ⊢ \fun{metisRule}~φ~ψ.
\end{equation*}

\end{myexamplenum}

The remainder of this section will be devoted to present a formal
description in type theory of remaining \Metis rules presented in
Table~\ref{tab:agda-metis-table} using the pattern presented in the
example above. We follow the same order to present the rules as the
table shows.

% ------------------------------------------------------------------
% Rules that follow the pattern:

\subfile{sections/proof-reconstruction/conjunct.tex}
\subfile{sections/proof-reconstruction/resolve.tex}
\subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}

% -------------------------------------------------------------------

We have finished the formalization of every inference rule in
a \Metis derivation, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena
jointly with the \Agda formalizations of the rules mentioned above
with more than eighty representative theorems in propositional
logic. An interested reader can test the
problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

\end{document}