### 2018-04-03 pubs/itp-2018/sections/acknowledgements.tex
[ pubs ] preparation of the paper (#60)
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict and also improved for readability.
Remove the numbers in the diagram.
Appendix updated removing the tstp of the small example.

* [ Sec4 ] example of the s

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

The first author would like to thank to Universidad EAFIT for funding
and supporting his master studies. He thankfully acknowledge Andreas
Abel and Chalmers University of Technology at Goteborg, Sweden for
inviting him to be part of the \Agda Implementors’ Meeting XXV where
he presented part of this research. We thank Joe Leslie-Hurd for his
support and comments about \Metis.  This research was fully supported
by Universidad EAFIT (grant \#~690-000088).

\end{document}