### 2018-04-03 pubs/itp-2018/sections/conclusions-and-future-work.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Conclusions}
\label{sec:conclusions}

We presented a proof-reconstruction approach in type theory for the
propositional fragment of the \Metis prover. We provided for each
\Metis inference rule a formal description in type theory following a
syntactical approach. This formalizations are mainly exposed in
Section~\ref{ssec:emulating-inferences}.

We built the \Athena translator tool written in \Haskell
that generates \Agda proof-terms of \Metis derivations.
\Agda files generated by this translator imports
\Agda formalizations of the \Metis reasoning~\cite{AgdaProp,AgdaMetis}.
% We tested successful using this tool and \Agda as we
% described in Section~\ref{sec:proof-reconstruction} over a subset of
% propositional problems from~\cite{Prieto-Cubides2017}.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{ssec:emulating-inferences}.
Nevertheless, we chose that syntactical treatment instead of using
semantics to extend this work towards the support of first-order
logic or other non-classical logics.
For first-order logic, recall satisfiability is undecidable and
its syntactical aspect plays an important role to reconstruct proofs.

One of the main contribution of this study was increasing the
trustworthiness of the automatic prover \Metis. Justifying a proof by
a theorem prover has a real significant impact for these automatic
tools.  The reverse engineering task to grasp the prover
reasoning can reveal important issues or bugs in many parts of these
systems (\eg, preprocessing, reasoning, or deduction modules). During
this research, we had the opportunity to contribute to \Metis by
reporting some bugs---see Issues No.~2, No.~4, and commit
\name{8a3f11e} in \Metis' official
repository.\footnote{\url{https://github.com/gilith/metis}.}
Fortunately, all these problems were fixed quickly by Hurd in
\Metis~2.3 (release~20170822).
\end{document}