### 2018-06-25 pubs/itp-2018/sections/acknowledgements.tex
Attempt to realocate the resolve rule after simplify

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

The first author would like to thank to Universidad EAFIT for funding
and supporting his master studies. We thank Joe Leslie-Hurd for his
support and comments about \Metis. This research was fully supported
by Universidad EAFIT (grant \#~690-000088).

\end{document}