### 2018-06-30 pubs/itp-2018/sections/acknowledgements.tex
First review of related work

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

The authors would like to thank to Universidad EAFIT for funding and supporting
this publication (grant \#~690-000088). We thank Joe Leslie-Hurd for his support
and comments about \Metis.

\end{document}