### 2018-07-25 pubs/itp-2018/sections/related-work.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

% ===================================================================

\section{Related Work}
\label{sec:related-work}

Many approaches have been proposed for proof-reconstruction and some tools have
been implemented in the last decades. We first mention some tools in type
theory and later we listed some proof-reconstruction tools for classical logic.

\citeauthor{Kanso2012} in~\cite{Kanso2012,kanso2016light} shows a
proof-reconstruction in \Agda using semantics for logic equivalences for
derivations in propositional logic generated by \name{EProver} and \name{Z3}. In
\cite{foster2011integrating}~\citeauthor{foster2011integrating}~describe another
proof-reconstruction in \Agda but for pure equational logic of derivations
generated by \name{Waldmeister}~\cite{hillenbrand1997}. No other
proof-reconstruction has been carried out neither in \Agda nor with \Metis
prover as far as we know.

In \name{Coq}~\cite{coqteam}, other proof-assistant closes to \Agda, we found
the \name{SMTCoq}~\cite{armand2011,Ekici2017} tool which provides a certified
checker for proof witness coming from the \SMT solver
\name{veriT}~\cite{bouton2009} and adds a new tactic named verit, that calls
\name{veriT} on any \name{Coq} goal. \citeauthor{Bezem2002} in~\cite{Bezem2002}
transform a proof produced by the first-order automatic theorem prover
\name{Bliksem}~\cite{deNivelle2003} in a \name{Coq} proof-term given a fixed but
arbitrary first-order signature.

There are other successful attempts to reconstruct proofs from \ATPs but
using proof-assistants for classical logic instead of type theory.
Let us mention some representative tools.

The \name{Isabelle} proof-assistant has the \name{Sledgehammer} tool.
This program provides a full integration between
automatic theorem provers~\cite{blanchette2013extending,Fleury2014,bohme2010} and
\name{Isabelle/HOL}~\cite{nipkow2002isabelle}, the specialization of
\name{Isabelle} for higher-order logic.
A modular proof-reconstruction workflow is presented jointly with
the full integration of \name{Leo-II} and \name{Satallax} provers with
\name{Isabelle/HOL} in \citeauthor{Een2004}~\cite{Een2004}.

\citeauthor{Hurd1999}~\cite{Hurd1999} integrates the first-order resolution
prover \name{Gandalf} with the high-order theorem prover
\name{HOL}~\cite{norrish2017hol}.
Its \verb!GANDALF_TAC! tactic is able to reconstruct \name{Gandalf} proofs
by using a \abbre{LCF} model. For \name{HOL Light}, a version of
\name{HOL} but with a simpler logic core, the \SMT solver \name{CVC4}
was integrated. \citeauthor{kaliszyk2013}~\cite{kaliszyk2013}
reconstruct proofs from different \ATPs with the \name{PRocH} tool by
replaying detailed inference steps from the \ATPs with internal
inference methods implemented in \name{HOL Light}.

\end{document}