### 2019-07-09 pubs/paper/sections/08-acknowledgements.tex

\documentclass[../paper.tex]{subfiles}
\begin{document}

% ===================================================================

\subsubsection*{Acknowledgments.}

The authors would like to thank to Universidad EAFIT for funding and
supporting this publication (grant \#~690-000088). We thank Joe
Leslie-Hurd for his support and comments about \Metis. 
% A mention of the agda-meeting where I received good feedback.
Special thanks to Juan Carlos Agudelo Agudelo and Håkon Robbestad
Gylterud, for reading and suggesting corrections of a preliminary
version of this documment.

\end{document}