### 2019-07-10 pubs/paper/sections/01-abstract.tex
PR (#68)
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict and also improved for readability.
Remove the numbers in the diagram.
Appendix updated removing the tstp of the small example.

* [ Sec4 ] example of the strip inference.

* First attemp

\documentclass[../paper.tex]{subfiles}
\begin{document}

\begin{abstract}

  We describe a syntactical proof-reconstruction approach to verifying
  derivations generated by the automatic theorem prover \Metis for
  classical, propositional logic. To verify the derivations, we use
  \Agda, a proof assistant for dependent type theory. We formalise in
  \Agda each inference rule of \Metis reasoning, that is, we transform
  \Metis derivations into \Agda proof-terms. In a future full-fledged
  integration of \Agda with off-the-shelf automatic theorem provers,
  proof-reconstruction shall be a necessary component.

  \keywords{~proof-reconstruction, automatic theorem provers,
    proof-assistants, Agda, Metis}
\end{abstract}

\end{document}